---
title: "Metadata"
author: "Joe Celebrezze"
date: "2023-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

# Raw Data
To be filled in later...

# Processed Data
## main_dataset.csv
This dataset is intended to be the primary dataset used in data analyses
```{r}
main_df <- read.csv(here('data', 'processed-data', 'main_dataset.csv'))
summary(main_df)
```

**species**: 6-letter species code w/ first three letters of genus name followed by first three letters of specific epiphet (exception: "MANDEN" which uses common name, 'Manzanita', as first three letters of code); see below for what species each code represents
    *Iris douglasiana* ~ "IRIDOU",
    *Erigeron karvinskianus* ~ "ERIKAR",
    *Salvia apiana* ~ "SALAPI",
    *Salvia leucophylla* ~ "SALLEU",
    *Artemesia californica* ~ "ARTCAL",
    *Ceanothus griseus* ~ "CEAGRI",
    *Malosma laurina* ~ "MALLAU",
    *Heteromeles arbutifolia* ~ "HETARB"
    *Arctostaphylos densiflora* ~ "MANDEN"
    
**plant**: numerical variable indicating which plant the sample came from (typically in between 1-12)

**date**: date, in format YYYY-MM-DD

**total_branch_mass**: total mass (g) of the branch used for each plant when calculating stem-to-leaf ratio

**total_leaf_mass**: total mass (g) of the leaves used for each plant when calculating stem-to-leaf ratio

**stem_mass_ratio**: % of total mass (leaves and stem) that is taken up by stem (one value per plant)

**leaf_mass_ratio**: % of total mass (leaves and stem) that is taken up by leaves (one value per plant)
  
**leaf_lfm**: live fuel moisture (%) of the leaves (one value per plant)

**thickness**: average leaf thickness (units?) as measured by calipers (one value per plant)

**leaf_wet_mass**: wet mass (g) of the leaves used in live fuel moisture measurement (typically five leaves, w/ ERIKAR and ARTCAL providing exceptions due to their very small leaves) (one value per plant)

**leaf_dry_mass**: dry mass (g) of the leaves used in live fuel moisture measurement (typically five leaves, w/ ERIKAR and ARTCAL providing exceptions due to their very small leaves) (one value per plant)

**stem_lfm**: live fuel moisture (%) of the stem (one value per plant)

**stem_width**: average width of the stem (units?); intended as a proxy for woodineess (? **confirm**) (one value per plant)

**stem_wet_mass**: wet mass (g) of the stem used in live fuel moisture measurement (one value per plant)

**stem_dry_mass**: dry mass (g) of the stem used in live fuel moisture measurement (one value per plant)

**lfm**: live fuel moisture of the plant (%, leaves and stem included) (one value per plant)

**leaf_area**: average area of five leaves measured by using ImageJ and leaf scans; used in the calculations of LMA and SLA below

**auc_voc**: area under the curve for the plot of VOC's over time; intended as a proxy for VOC emmissions (*one value per species*)

**avg_voc**: average concentration of VOC's as detected by the VOC meter utilized by Jackie Liang (VOCs/second) (one value per plant) (*one value per species*)

**LMA**: leaf mass per area (g/cm^2) (note: NAs for ERIKAR and ARTCAL due to those species not having the correct data to be able to calculate LMA) (one value per plant)

**SLA**: specific leaf area (cm^2/g) (note: NAs for ERIKAR and ARTCAL for same reason as above) (one value per plant)

**rep**: numerical indicator of the sample used in the flammability test; typically, each plant was burned multiple times

**branch_length**: length (cm) of the sample loaded into the flammability chamber

**branch_width**: width (cm) of the sample loaded into the flammability chamber

**branch_height**: : height (cm) of the sample loaded into the flammability chamber

**sample_wt**: weight (g) of the sample loaded into the flammability chamber

**branching**: indicator of the number of branches on the sample loaded into the flammability chamber; showing sample morphology, 'bushiness', etc.                    

**mpa**: water potential (-MPa)

**notes_on_plant_char**: notes on plant characteristics, as taken during the flammability testing

**start_time**: time that the flammability testing began for given sample

**ambient_temp**: ambient temp (degrees C) as taken by small sensor on front of flammability chamber; note: this was influenced by the temperature of the chamber and therefore is not entirely reliable

**ambient_humidity**: ambient humidity (%) as taken by small sensor on front of flammability chamber; note: this was influenced by the temperature of the chamber and therefore is not entirely reliable

**pre_ignition_glow**: binary indicator (1 or 0) of whether or not the sample glowed prior to ignition

**first_glow_time**: the time that the sample started glowing first, if visible

**ignition**: indicator of whether the sample ignited: 1 = ignited on its own, M = manual ignition, 0 = no ignition
  
**primary_ignition**: the time that the sample ignited for the first time, if applicable

**primary_time_of_flame_end**: the time that the first flame ended, if applicable

**secondary_ignition**: the time that the sample ignited for the second time, if applicable

**secondary_time_of_flame_end**: the time that the second flame ended, if applicable

**third_ignition**: the time that the sample ignited for the third time, if applicable (a third flame only happened a few times)

**third_time_of_ignition_flame_end**: the time that the third flame ended, if applicable (a third flame only happened a few times)

**time_fh**: the time that the sample reached it's highest flame height

**fh**: maximum flame height (cm)

**time_of_glow_end**: the time that glowing ended (note: in many cases, this was unclear on the videos, so it was indicated during flammability testing by clapping; however, the 'clap' was not always right when the sample stopped glowing and, therefore, the glow metrics are somewhat unreliable)

**end_time**: the time that the test ended

**thermocoupler_height**: the height of the mobile thermocoupler in the back of the chamber

**hotplate_height**: the height of the hot plate, which was shifted up and down based on branch_height and species differences so that the sample was not scorched by the pilot flame, but it was sufficiently close for the pilot flame to adequately ignite the sample

**notes_on_flam_data**: notes on flammability data, taken during flammability testing and during video analysis

**fd**: flame duration (seconds); for cases where the sample burned multiple times, the most prominent flame was selected for flame duration (typically the flame with the largest flame height and the longest flame duration)

**tti**: time from the start of the test to the first ignition (seconds)

**pfg**: post-flame glow (seconds); see note on glow above

**max_temp**: maximum temperature (degrees C) in the time frame starting w/ primary_ignition and ending with primary_time_of_flame_end

**time_at_max_temp**: the time that the maximum temperature occurred

**max_temp_sensor**: the thermocoupler sensor used in calculating the maximum temperature

**start_temp**: the maximum starting temperature (across all thermocouplers, degrees C)

**start_temp_sensor**: the thermocoupler sensor used in calculating the starting temperature

**temp_change**: the change in temperature from start to maximum (degrees C)

**max_heat_flux_loessCH7**: maximum heat flux (using CH7 sensor) as calculated using fitted values using Loess regressions w/ an optimal smoothing span as derived from an optimization routine

**max_heat_flux_loessCH8**: maximum heat flux (using CH8 sensor) as calculated using fitted values using Loess regressions w/ an optimal smoothing span as derived from an optimization routine

**time_at_max_heat_flux_loessCH7**: time of maximum heat flux (using CH7 sensor)

**time_at_max_heat_flux_loessCH8**: time of maximum heat flux (using CH8 sensor)

**avg_heat_flux_stableCH7**: average heat flux (CH7 sensor) during the 'stable' period of time (15 seconds after start time, 10 seconds prior to ignition time)

**avg_heat_flux_stableCH8**: average heat flux (CH8 sensor) during the 'stable' period of time (15 seconds after start time, 10 seconds prior to ignition time)

**heat_flux_change**: intended to be the main heat flux variable in the analyses; calculated as: max_heat_flux_loessCH8 - avg_heat_flux_stableCH8; CH8 was used instead of CH7 due to a higher signal-to-noise ratio

**prop_ig**: % of samples that ignited naturally (*one value per species*)

**wet_mass**: the wet mass of the sample used in calculating live fuel moisture (one value per plant); used to calculate dw_flam_sample and ww_flam_sample

**dry_mass**: the dry mass of the sample used in calculating live fuel moisture (one value per plant); used to calculate dw_flam_sample and ww_flam_sample

**gdw_gfw**: equal to dry_mass/(wet_mass+dry_mass); used to calculate dw_flam_sample and ww_flam_sample

**dw_flam_sample**: the dry mass of the sample (g), as calculated using live fuel moisture and the sample_wt

**ww_flam_sample**: the wet mass of the sample (g), as calculated using live fuel moisture and the sample_wt

**branch_volume**: the sample volume (cm^3), as calculated using branch_height, branch_width, and branch_length