---
title: "Species Differences, Looking at Species-Specific Variables, Average Values
  for Species"
author: "Joe Celebrezze"
date: "2023-07-17"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
here = here::here
source(here("scripts", "source_code.R"))
```

# Data Wrangling
```{r}
flamm.df <- flamm.df %>% 
  mutate(species = as.factor(species)) %>% 
  unite('species_id', c(species, plant), sep = '_', remove = F) # adding species_id for random effects
```

# Summary Tables
```{r}
flamm.df %>% 
  select(species, sample_wt) %>% 
  group_by(species) %>% 
  dplyr::summarise(sw = mean(sample_wt))

flamm.df %>% 
  drop_na(fd, fh, temp_change, heat_flux_change) %>% 
  group_by(species) %>% 
  dplyr::summarise(fd = mean(fd), fh = mean(fh),
                   temp_change = mean(temp_change), heat_flux_change = mean(heat_flux_change))
```


Although the mixed effects model selection tells us a lot about our data and provides useful insights, it doesn't include all of the variables and it leaves some of our question unanswered:

*Question*: How can interspecific differences in flammability be described by plant traits -- plant hydration, plant morphology, leaf traits, VOC's?

With the mixed effects models, we looked into which variables were key in predicting the flammability metrics, but we did not necessarily look into how the above question fully. To do so, let's look at interspecific differences in flammability metrics and then interspecific differences in traits of interest and compare them.

We are also going to do the k-means cluster analysis here on a full dataset, a dataset with flammability metrics, and a dataset with traits to try and get at the question above

# Boxplots, ANOVAs

## Flammability Metrics

### Flame Height
```{r}
fh.spp <- lmer(fh ~ species + (1|species_id), data = flamm.df)
anova(fh.spp) # p < 0.00001
fh.spp.tukey <- emmeans(fh.spp, pairwise ~ species, adjust = 'tukey')
fh.spp.tukey.df <- summary(fh.spp.tukey$contrasts)
fh.spp.tukey.labels <- data.frame(species = c('HETARB', 'CEAGRI', 'SALAPI', 'MALLAU', 'SALLEU', 'ARCDEN', 'ARTCAL', 'ERIKAR'), letters = c('a', 'ab', 'ab', 'ab', 'ab', 'abc', 'bc', 'c'), adj = c(0, 0, 4, 4, 0, 4, 0, 4)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

fh.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU', 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = fh, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = fh, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 65 + adj, label = letters), fontface = 'bold',
              data = fh.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Flame Height (cm)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
fh.boxplot
```

#### Graphical Abstract Version
```{r}
flamm.df %>%
  filter(species %in% c('CEAGRI', 'SALAPI', 'ARTCAL', 'ERIKAR')) %>% 
  mutate(species = case_when(species == 'CEAGRI' ~ 'C. griseus',
                             species == 'SALAPI' ~ 'S. apiana',
                             species == 'ARTCAL' ~ 'A. californica',
                             species == 'ERIKAR' ~ 'E. karvinskianus')) %>% 
  mutate(species = fct_relevel(species, c('C. griseus', 'S. apiana',
                                          'A. californica', 'E. karvinskianus'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = fh), outlier.shape = NA, alpha= 0.8, fill = 'gray80') +
    geom_jitter(aes(x = species, y = fh), width = 0.2, alpha = 0.2) +
    labs(x = 'Species', y = 'Flame Height (cm)') +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 24),
          axis.title.x = element_blank(),
          axis.text.x = element_text(face = 'bold.italic', size = 22,
                                     angle = 15, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 18))
```


### Flame Duration
```{r}
fd.spp <- lmer(fd ~ species + (1|species_id), data = flamm.df)
anova(fd.spp) # p < 0.00001
fd.spp.tukey <- emmeans(fd.spp, pairwise ~ species, adjust = 'tukey')
fd.spp.tukey.df <- summary(fd.spp.tukey$contrasts)
fd.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'MALLAU', 'HETARB', 'SALLEU', 'ARCDEN', 'SALAPI', 'ARTCAL', 'ERIKAR'), letters = c('ab', 'a', 'ab', 'abc', 'abcd', 'bcd', 'cd', 'd'), adj = c(0, 4, 0, 0, 4, 4, 0, 4)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

fd.boxplot <- flamm.df %>%
  filter(fd < 70) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = fd, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = fd, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 59.5 + adj, label = letters), fontface = 'bold',
              data = fd.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Flame Duration (sec)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
fd.boxplot
```

### Temp. Change
```{r}
tc.spp <- lmer(temp_change ~ species + (1|species_id), data = flamm.df)
anova(tc.spp) # p < 0.00001
tc.spp.tukey <- emmeans(tc.spp, pairwise ~ species, adjust = 'tukey')
tc.spp.tukey.df <- summary(tc.spp.tukey$contrasts)
tc.spp.tukey.labels <- data.frame(species = c('HETARB', 'MALLAU', 'SALLEU', 'SALAPI', 'CEAGRI', 'ARCDEN', 'ARTCAL', 'ERIKAR'), letters = c('a', 'ab', 'ab', 'ab', 'ab', 'ab', 'b', 'b'),
                                  adj = c(0, 3, 0, 3, 0, 3, 0, 3)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

tc.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  filter(temp_change < 100) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = temp_change, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = temp_change, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 38 + adj, label = letters), fontface = 'bold',
              data = tc.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Temp. Change (°C)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
tc.boxplot
```

### Heat Flux Change
```{r}
hf.spp <- lmer(heat_flux_change ~ species + (1|species_id), data = flamm.df)
anova(hf.spp) # p < 0.00001
hf.spp.tukey <- emmeans(hf.spp, pairwise ~ species, adjust = 'tukey')
hf.spp.tukey.df <- summary(hf.spp.tukey$contrasts)
hf.spp.tukey.labels <- data.frame(species = c('HETARB', 'SALAPI', 'MALLAU', 'CEAGRI', 'SALLEU', 'ARCDEN', 'ARTCAL', 'ERIKAR'), letters = c('a', 'ab', 'ab', 'abc', 'bc', 'bc', 'bc', 'c'), adj = c(0, 100, 100, 0, 0, 100, 0, 100)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

hf.boxplot <- flamm.df %>%
  #filter(heat_flux_change < 0.3) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = heat_flux_change, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = heat_flux_change, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 6000 + adj, label = letters), fontface = 'bold',
              data = hf.spp.tukey.labels, size = 6) +
    #scale_y_continuous(breaks = c(0.05, 0.1, 0.15, 0.20)) +
    labs(x = 'Species', y = expression(bold("Heat Flux "(W/m^2)))) +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.y = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
hf.boxplot
```

## Traits

### Sample Weight
Because all of the 'sample size' metrics are pretty similar, I think looking at only one, at least for now will work. Wet weight has been instrumental in previous analyses, so we will use it below
```{r}
sw.spp <- lmer(sample_wt ~ species + (1|species_id), data = flamm.df)
anova(sw.spp) # p < 0.00001
sw.spp.tukey <- emmeans(sw.spp, pairwise ~ species, adjust = 'tukey')
sw.spp.tukey.df <- summary(sw.spp.tukey$contrasts)
sw.spp.tukey.labels <- data.frame(species = c('HETARB', 'MALLAU', 'SALAPI', 'CEAGRI', 'SALLEU', 'ARCDEN',  'ARTCAL', 'ERIKAR'), letters = c('a', 'ab', 'bc', 'cd', 'd', 'cd', 'd', 'd'), 
                                  adj = c(0, 1.7, 1.7, 0, 0, 1.7, 0, 1.7)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

sw.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = sample_wt, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = sample_wt, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 24.6 + adj, label = letters), fontface = 'bold',
              data = sw.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Sample Weight (g)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
sw.boxplot
```

### MPa
```{r}
mpa.spp <- lmer(mpa ~ species + (1|species_id), data = flamm.df)
anova(mpa.spp) # p < 0.00001
mpa.spp.tukey <- emmeans(mpa.spp, pairwise ~ species, adjust = 'tukey')
mpa.spp.tukey.df <- summary(mpa.spp.tukey$contrasts)
mpa.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'ARTCAL', 'ERIKAR', 'HETARB', 'MALLAU', 'SALLEU', 'ARCDEN', 'SALAPI'), letters = c('a', 'b', 'bc', 'bc', 'bc', 'bc', 'bc', 'c'), adj = c(0, 0, 0.025, 0, 0.025, 0, 0.025, 0.025)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

mpa.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = mpa, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = mpa, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 0.41 + adj, label = letters), fontface = 'bold',
              data = mpa.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Water Potential (-MPa)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
mpa.boxplot
```

### Branching
```{r}
branching.spp <- lmer(branching ~ species + (1|species_id), data = flamm.df)
anova(branching.spp) # p < 0.00001
branching.spp.tukey <- emmeans(branching.spp, pairwise ~ species, adjust = 'tukey')
branching.spp.tukey.df <- summary(branching.spp.tukey$contrasts)
branching.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'ARCDEN', 'SALLEU', 'ARTCAL', 'MALLAU', 'HETARB', 'ERIKAR', 'SALAPI'), letters = c('a', 'ab', 'abc', 'ab', 'abcd', 'bcd', 'd', 'cd'),
                                         adj = c(0, 0.035, 0, 0, 0.035, 0, 0.035, 0.035)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

branching.boxplot <- flamm.df %>%
  filter(branching < 0.75) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = branching, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = branching, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 0.65 + adj, label = letters), fontface = 'bold',
              data = branching.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Branching') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
branching.boxplot
```

### Starting Temp
```{r}
st.spp <- lmer(start_temp ~ species + (1|species_id), data = flamm.df)
anova(st.spp) # p < 0.00001
st.spp.tukey <- emmeans(st.spp, pairwise ~ species, adjust = 'tukey')
st.spp.tukey.df <- summary(st.spp.tukey$contrasts)
st.spp.tukey.labels <- data.frame(species = c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                              'ARCDEN', 'ARTCAL', 'ERIKAR'),
                                  letters = c('b', 'b', 'ab', 'b', 'b', 'ab', 'a', 'ab'),
                                  adj = c(0, 6, 0, 6, 0, 6, 0, 6)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)


flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = start_temp, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = start_temp, color = species), width = 0.2, alpha = 0.2) +
    labs(x = 'Species', y = 'Start Temp.') +
    geom_text(aes(x = species, y = 240 + adj, label = letters), fontface = 'bold',
              data = st.spp.tukey.labels, size = 6) +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
```

### Dotplot Data Wrangling
```{r}
flamm.dotplot.df1 <- flamm.df %>% 
  group_by(species, plant) %>% 
  summarise(lfm = mean(lfm), dmc = mean(dmc), leaf_mass_ratio = mean(leaf_mass_ratio))

flamm.dotplot.df2 <- flamm.dotplot.df1 %>% 
  group_by(species) %>% 
  summarise(sd_lfm = sd(lfm), sd_dmc = sd(dmc), sd_leaf_mass_ratio = sd(leaf_mass_ratio),
            lfm = mean(lfm), dmc = mean(dmc), leaf_mass_ratio = mean(leaf_mass_ratio))
```


### LFM
Boxplot
```{r}
lfm.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = lfm, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = lfm, color = species), width = 0.2, alpha = 0.2) +
    labs(x = 'Species', y = 'Live Fuel Moisture (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
lfm.boxplot
```

Dotplot
```{r}
lfm.dotplot <- flamm.dotplot.df2 %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, color = species)) +
    geom_point(aes(y = lfm), size = 10) +
    geom_linerange(aes(ymin = lfm-sd_lfm, ymax = lfm+sd_lfm), linewidth = 1) +
    labs(x = 'Species', y = 'Live Fuel Moisture (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
lfm.dotplot
```


### DMC
Boxplot
```{r}
dmc.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = dmc, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = dmc, color = species), width = 0.2, alpha = 0.2) +
    labs(x = 'Species', y = 'Dry Matter Content (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
dmc.boxplot
```

Dotplot
```{r}
dmc.dotplot <- flamm.dotplot.df2 %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, color = species)) +
    geom_point(aes(y = dmc), size = 10) +
    geom_linerange(aes(ymin = dmc-sd_dmc, ymax = dmc+sd_dmc), linewidth = 1) +
    labs(x = 'Species', y = 'Dry Matter Content (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
dmc.dotplot
```


### Leaf Mass Ratio
Boxplot
```{r}
lmr.boxplot <- flamm.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = leaf_mass_ratio, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = leaf_mass_ratio, color = species), width = 0.2, alpha = 0.2) +
    labs(x = 'Species', y = 'Leaf Mass Ratio (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
lmr.boxplot
```

Dotplot
```{r}
lmr.dotplot <- flamm.dotplot.df2 %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  mutate(sd_leaf_mass_ratio = sd_leaf_mass_ratio*100, leaf_mass_ratio = leaf_mass_ratio*100) %>% 
  ggplot(aes(x = species, color = species)) +
    geom_point(aes(y = leaf_mass_ratio), size = 10) +
    geom_linerange(aes(ymin = leaf_mass_ratio-sd_leaf_mass_ratio, ymax = leaf_mass_ratio+sd_leaf_mass_ratio), linewidth = 1) +
    labs(x = 'Species', y = 'Leaf Mass Ratio (%)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
lmr.dotplot
```


## Arranging
```{r}
# blank figure
blank <- ggplot(data = flamm.df, aes(x = species, y = mpa)) +
  theme_void()
blank

# abridged version
top <- ggarrange(fh.boxplot, fd.boxplot, hf.boxplot,
          labels = c('A', 'B', 'C'),
          font.label = list(size = 28), ncol = 3, nrow = 1)
top

bottom <- ggarrange(blank, tc.boxplot, sw.boxplot, blank,
          labels = c('', 'D', 'E', ''),
          font.label = list(size = 28),
          widths = c(0.48, 1, 1, 0.48), ncol = 4, nrow = 1)
bottom

ggarrange(top, bottom, nrow = 2, ncol = 1)
#ggsave(here('figures', 'main_figures', 'five.panel.spp.boxplots.png'), height = 9, width = 13.5) # height was 9 w/ two rows

# expanded version
ggarrange(fh.boxplot, fd.boxplot, hf.boxplot, tc.boxplot, sw.boxplot, branching.boxplot, lfm.dotplot, dmc.dotplot, lmr.dotplot,
          labels = c('A', 'B', 'C', 'D', 'E', 'F', 'G', 'H', 'I', 'J'),
          font.label = list(size = 28), ncol = 3, nrow = 3)


#ggsave(here('figures', 'main_figures', 'spp.boxplots.png'), height = 14, width = 13.5) # height was 9 w/ two rows

# dotplots
spp.dotplots <- ggarrange(lfm.dotplot, dmc.dotplot, lmr.dotplot, ncol = 3, nrow = 1)
spp.dotplots

ggsave(here('figures', 'extra_figures', 'spp.dotplots.png'), height = 4, width = 13.5) # height was 9 w/ two rows
```

# Main Figure: Boxplots
## Data Wrangling
Tukey-Kramer Results
```{r}
fh.spp.tukey.labels <- fh.spp.tukey.labels %>% 
  mutate(variable = 'fh') %>% 
  mutate(y.pos = adj + 65)
fd.spp.tukey.labels <- fd.spp.tukey.labels %>% 
  mutate(variable = 'fd') %>% 
  mutate(y.pos = adj + 59.5)
tc.spp.tukey.labels <- tc.spp.tukey.labels %>% 
  mutate(variable = 'temp_change') %>% 
  mutate(y.pos = adj + 38)
hf.spp.tukey.labels <- hf.spp.tukey.labels %>% 
  mutate(variable = 'heat_flux_change') %>% 
  mutate(y.pos = adj + 6000)
sw.spp.tukey.labels <- sw.spp.tukey.labels %>% 
  mutate(variable = 'sample_wt') %>% 
  mutate(y.pos = adj + 24.6)
branching.spp.tukey.labels <- branching.spp.tukey.labels %>% 
  mutate(variable = 'branching') %>% 
  mutate(y.pos = adj + 0.65)
# if you prefer boxplots of LFM, DMC, LMR, then include this in tukey.labels.df
no.tukey.labels.df <- data.frame(species = c(rep(c("HETARB", "CEAGRI", "SALAPI", "MALLAU", "SALLEU", "ARCDEN", "ARTCAL", "ERIKAR"), 3)),
                                 letters = c(rep('', 24)),
                                 adj = c(rep(0, 24)),
                                 y.pos = c(rep(0, 24)),
                                 variable = c(rep('lfm', 8), rep('dmc', 8),
                                              rep('leaf_mass_ratio', 8)))
tukey.labels.df <- rbind(fh.spp.tukey.labels, fd.spp.tukey.labels, tc.spp.tukey.labels, hf.spp.tukey.labels, sw.spp.tukey.labels, branching.spp.tukey.labels)
```

## Visualization
```{r}
variable_names <- c(fh = 'Flame Height (cm)',
                    fd = 'Flame Duration (sec)',
                    temp_change = 'Temp. Change (°C)',
                    heat_flux_change = 'Heat Flux Change (W/m2)',
                    sample_wt = 'Sample Mass (g)',
                    branching  = 'Branching (branches/cm)',
                    lfm  = 'Live Fuel Moisture (%)',
                    dmc  = 'Dry Matter Content (%)',
                    leaf_mass_ratio = 'Leaf Mass Ratio (%)')

main.boxplots <- flamm.df %>% 
  filter(fd < 70, temp_change < 100, branching < 0.75, leaf_mass_ratio > 0.4) %>% 
  pivot_longer(cols = c(fh, fd, temp_change, heat_flux_change, sample_wt,
                        branching),
               names_to = 'variable',
               values_to = 'value') %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU', 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = value, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = value, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = y.pos, label = letters), fontface = 'bold',
              data = tukey.labels.df, size = 6) +
    facet_wrap(~factor(variable, c('fh', 'fd', 'heat_flux_change', 'temp_change', 'sample_wt',
                        'branching')),
               scales = 'free_y',
               strip.position = 'left',
               labeller = as_labeller(variable_names)) +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          #axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75), # hash out if including dotplots
          axis.text.y = element_text(size = 15),
          strip.background = element_blank(),
          strip.placement = 'outside',
          strip.text = element_text(size = 17, face = 'bold'))
main.boxplots

plot_grid(main.boxplots, spp.dotplots, rel_heights = c(2, 1.12), ncol = 1)

ggsave(here('figures', 'main_figures', 'clean.spp.boxplots.png'), height = 12, width = 13.5) # height was 9 w/ two rows
```


# REMEF Boxplots, ANOVAs
```{r}
interspec.df <- read.csv(here('data', 'processed-data', 'interspecific_flam_remef.csv')) %>% 
    unite('species_id', c(species, plant), remove = F)
```

## Flame Height
```{r}
fh.spp <- lmer(r_fh ~ species + (1|species_id), data = interspec.df)
anova(fh.spp) # p < 0.00001
fh.spp.tukey <- emmeans(fh.spp, pairwise ~ species, adjust = 'tukey')
fh.spp.tukey.df <- summary(fh.spp.tukey$contrasts)
fh.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'SALLEU', 'ARTCAL', 'SALAPI', 'ARCDEN', 'HETARB', 'MALLAU', 'ERIKAR'), letters = c('a', 'ab', 'ab', 'ab', 'ab', 'abc', 'bc', 'c'), adj = c(0, 0, 0, 0, 0, 0, 0, 0)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

fh.boxplot <- interspec.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU','ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = r_fh, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = r_fh, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 1.3 + adj, label = letters), fontface = 'bold',
              data = fh.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Flame Height (cm)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
fh.boxplot
```

## Flame Duration
```{r}
fd.spp <- lmer(r_fd ~ species + (1|species_id), data = interspec.df)
anova(fd.spp) # p < 0.00001
fd.spp.tukey <- emmeans(fd.spp, pairwise ~ species, adjust = 'tukey')
fd.spp.tukey.df <- summary(fd.spp.tukey$contrasts)
fd.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'SALLEU', 'ARTCAL', 'SALAPI', 'ARCDEN', 'HETARB', 'MALLAU', 'ERIKAR'), letters = c('a', 'ab', 'abc', 'c', 'abc', 'abc', 'abc', 'bc'), adj = c(0, 0, 0, 0, 0, 0, 0, 0)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

fd.boxplot <- interspec.df %>%
  filter(fd < 70) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU','ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = r_fd, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = r_fd, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 4.25 + adj, label = letters), fontface = 'bold',
              data = fd.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Flame Duration (sec)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
fd.boxplot
```

## Temp. Change
```{r}
tc.spp <- lmer(r_temp_change ~ species + (1|species_id), data = interspec.df)
anova(tc.spp) # p < 0.00001
tc.spp.tukey <- emmeans(tc.spp, pairwise ~ species, adjust = 'tukey')
tc.spp.tukey.df <- summary(tc.spp.tukey$contrasts)
tc.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'SALLEU', 'ARTCAL', 'SALAPI', 'ARCDEN', 'HETARB', 'MALLAU', 'ERIKAR'), letters = c('a', 'a', 'a', 'a', 'a', 'a', 'a', 'a'),
                                  adj = c(0, 0, 0, 0, 0, 0, 0, 0)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

tc.boxplot <- interspec.df %>%
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU','ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  filter(temp_change < 100) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = r_temp_change, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = r_temp_change, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 1.5 + adj, label = letters), fontface = 'bold',
              data = tc.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Temp. Change (°C)') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
tc.boxplot
```

## Heat Flux Change
```{r}
hf.spp <- lmer(r_heat_flux_change ~ species + (1|species_id), data = interspec.df)
anova(hf.spp) # p < 0.00001
hf.spp.tukey <- emmeans(hf.spp, pairwise ~ species, adjust = 'tukey')
hf.spp.tukey.df <- summary(hf.spp.tukey$contrasts)
hf.spp.tukey.labels <- data.frame(species = c('CEAGRI', 'SALLEU', 'ARTCAL', 'SALAPI', 'ARCDEN', 'HETARB', 'MALLAU', 'ERIKAR'), letters = c('a', 'ab', 'ab', 'a', 'ab', 'ab', 'b', 'ab'), adj = c(0, 0, 0, 0, 0, 0, 0, 0)) # manually extracted from above dataframe (couldn't figure out way to automate process w/ emmeans)

hf.boxplot <- interspec.df %>%
  filter(heat_flux_change < 0.3) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU','ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot() +
    geom_boxplot(aes(x = species, y = r_heat_flux_change, color = species), outlier.shape = NA, alpha= 0.8) +
    geom_jitter(aes(x = species, y = r_heat_flux_change, color = species), width = 0.2, alpha = 0.2) +
    geom_text(aes(x = species, y = 0.9 + adj, label = letters), fontface = 'bold',
              data = hf.spp.tukey.labels, size = 6) +
    labs(x = 'Species', y = 'Heat Flux Change') +
    scale_color_manual(values = c('#925E9FB2', '#AD002AB2', '#42B540B2', '#ADB6B6B2', '#1B1919B2','#00468BB2', '#ED0000B2', '#0099B4B2')) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 20, vjust = 0.9, hjust=0.75),
          axis.text.y = element_text(size = 15))
hf.boxplot
```

## Arranged
```{r}
ggarrange(fh.boxplot, fd.boxplot, hf.boxplot, tc.boxplot,
          labels = c('A', 'B', 'C', 'D'),
          font.label = list(size = 28), ncol = 2, nrow = 2)
#ggsave(here('figures', 'main_figures', 'remef.spp.boxplots.png'), height = 9.5, width = 9)
```

# ---------------------------------

# K-means Clustering
This analysis was not retained in the main text of the analysis, but it was included in the supplementary index and could be builded upon in future research

## Flamm Metrics
NOTE: Ignition method has significant effect on clustering
```{r}
id <- c(1:nrow(flamm.df))
flamm.df <- cbind(flamm.df, id) %>% 
  unite('species_id', c(species, id), remove = F)

# so it's replicable
set.seed(12)

# labels
flam_labels <- c('Flame Height', 'Flame Duration', 'Temperature Change', 'Heat Flux Change', '% Ignited')
names(flam_labels) <- c('fh', 'fd', 'temp_change', 'heat_flux_change', 'prop_ig')

# prepping dataset
flamm_kmeans_df <- flamm.df %>% 
  filter(ignition != 0) %>% 
  filter(temp_change < 100) %>% 
  select(species_id, species, ignition, fh, fd, temp_change, heat_flux_change, prop_ig) %>% 
  na.omit()

rownames(flamm_kmeans_df) <- flamm_kmeans_df$species_id

flamm_kmeans_scaled_df <- flamm_kmeans_df %>% 
  select(-c(species_id, species, ignition)) %>%
  scale() %>% 
  as.data.frame()

# optimal number of clusters
# 2 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
fviz_nbclust(flamm_kmeans_scaled_df, kmeans, method = "wss") # k = 3 or k = 4
# 2. silhouette method
fviz_nbclust(flamm_kmeans_scaled_df, kmeans, method = "silhouette") #k = 2

# run k-means clustering
#perform k-means clustering with k = 3 clusters
km <- kmeans(flamm_kmeans_scaled_df, centers = 3)
km$cluster <- case_when(km$cluster == 1 ~ 2,
                         km$cluster == 2 ~ 1,
                         km$cluster == 3 ~ 3)
#plot results of final k-means model
fviz_cluster(km, data = flamm_kmeans_scaled_df)
#add cluster assigment to original data
flamm_kmeans_df <- cbind(flamm_kmeans_df, cluster = km$cluster)
#merging with dataseet w/ plant traits (flamm.df) to visualize how clusters track plant traits
comp_kmeans_df <- merge(flamm_kmeans_df, flamm.df, by = 'species_id')

#visualizing how cluster assignment tracks species
flamm_kmeans_df %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    scale_y_continuous(breaks = c(1:max(flamm_kmeans_df$cluster))) +
    theme_bw() +
    labs(x = ' ', y = 'Cluster #', color = 'Cluster #', title = 'K-Means Clustering: Flam. Metrics')

flamm_kmeans_df %>% 
  ggplot(aes(x = cluster, fill = species)) +
    geom_bar() +
    theme_bw() +
    labs(x = 'Cluster #', fill = 'Species', y = 'Count', title = 'K-Means Clustering: Flam. Metrics')

flamm_kmeans_df %>% 
  pivot_longer(cols = c(fd, fh, heat_flux_change, temp_change, prop_ig), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(cluster), y = val, color = as.factor(cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free', labeller = labeller(var = flam_labels)) +
    labs(x = 'K-Means Cluster', y = 'Flamm. Metric Value') +
    scale_color_manual(values = c("#00468BB2", 'black', "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 17),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))

flamm_kmeans_df %>% 
  group_by(cluster, species) %>% 
  tally() %>%  
  pivot_wider(names_from = cluster, values_from = n) %>% 
  mutate(`1` = ifelse(is.na(`1`), 0, `1`),
         `2` = ifelse(is.na(`2`), 0, `2`)) %>% 
  mutate(kmeans.flam.index = (`2`)/(`2` + `1`)) %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = kmeans.flam.index, color = species)) +
    geom_point(size = 7) +
    labs(x = 'Species', y = 'K-Means Clustering: Combustibility Index') +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 14),
          axis.text = element_text(size = 13),
          axis.text.x = element_text(angle = 15))

flamm_kmeans_df %>% 
  group_by(ignition, cluster) %>% 
  tally()

ignition.vs.cluster <- lm(cluster ~ ignition, data = flamm_kmeans_df)
anova(ignition.vs.cluster)
summary(ignition.vs.cluster)

# how do plant traits relate to flam. clusters?
comp_kmeans_df %>% 
  pivot_longer(cols = c(lfm, mpa, dmc, sample_wt), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(cluster), y = val, color = as.factor(cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free') +
    labs(x = 'K-Means Cluster', y = 'Predictor Value') +
    scale_color_manual(values = c("#00468BB2", 'black', "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 17),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))
```

### Nice Figures
Alternative boxplot and reworking above figures to arrange:
```{r}
km_boxplot_flam <- comp_kmeans_df %>% 
  mutate(fh = scale(fh.x), fd = scale(fd.x), temp_change = scale(temp_change.x), heat_flux_change = scale(heat_flux_change.x), prop_ig = scale(prop_ig.x), lfm = scale(lfm), branching = scale(branching), leaf_sav = scale(leaf_sav), dmc = scale(dmc), sample_wt = scale(sample_wt)) %>% 
  pivot_longer(cols = c(fh, fd, temp_change, heat_flux_change, prop_ig, lfm, branching, leaf_sav, dmc, sample_wt), names_to = 'var', values_to = 'val') %>% 
  mutate(var = case_when(
    var == 'fh' ~ 'Flame Height',
    var == 'fd' ~ 'Flame Duration',
    var == 'heat_flux_change' ~ 'Heat Flux Change',
    var == 'temp_change' ~ 'Temp. Change',
    var == 'prop_ig' ~ 'Proportion Ignited',
    var == 'lfm' ~ 'Live Fuel Moisture',
    var == 'sample_wt' ~ 'Sample Weight',
    var == 'dmc' ~ 'Dry Mass Content',
    var == 'branching' ~ 'Branching',
    var == 'leaf_sav' ~ 'Leaf SA:Volume')) %>%
    mutate(var = fct_relevel(var,  c('Flame Height', 'Flame Duration', 'Heat Flux Change',
                                     'Temp. Change', 'Proportion Ignited',
                                     'Live Fuel Moisture', 'Sample Weight',
                                     'Dry Mass Content', 'Branching',
                                     'Leaf Surface Area:Volume'))) %>% 
  ggplot(aes(x = var, y = val, color = as.factor(cluster))) +
    geom_boxplot(alpha= 0.8) +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.text.x = element_text(face = 'bold', size = 15, angle = 20, vjust = 0.9, hjust=0.75),
          axis.title = element_blank(),
          axis.text.y = element_text(size = 15))
km_boxplot_flam

# Species vs. clusters
number_pts_per_cluster <- comp_kmeans_df %>% 
  dplyr::group_by(species.x, cluster) %>% 
  tally()

spp_vs_flamkm <- comp_kmeans_df %>% 
  mutate(species = fct_relevel(species.x, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    geom_text(aes(x = species.x, y = cluster + 0.25, color = as.factor(cluster), label = n),
                  data = number_pts_per_cluster, fontface = 'bold', size = 5.5) +
    annotate('text', x = 0.55, y = 1, label = 'High', angle = 90, fontface = 'bold',
             size = 5.5, color = "#00468BB2") +
    annotate('text', x = 0.95, y = 2, label = 'Combustibility/\nSustainability', angle = 90, fontface = 'bold',
             size = 5.5, color = "#ED0000B2") +
    annotate('text', x = 0.55, y = 3, label = 'Low', angle = 90, fontface = 'bold',
             size = 5.5, color = "#42B540B2") +
    annotate('text', x = 9, y = 2, label = 'Most Ignitable', angle = 270, fontface = 'bold',
             size = 5.5, color = "#ED0000B2") +
    geom_segment(aes(x = 0.55, y = 1.6, xend = 0.55, yend = 1.2),
                  arrow = arrow(length = unit(0.2, "cm")), color = "#00468BB2") +
    geom_segment(aes(x = 0.55, y = 1.6, xend = 0.55, yend = 2.4), color = "#ED0000B2") +
    geom_segment(aes(x = 0.55, y = 2.4, xend = 0.55, yend = 2.8),
                  arrow = arrow(length = unit(0.2, "cm")), color = "#42B540B2") +
    scale_y_continuous(breaks = c(1:max(comp_kmeans_df$cluster))) +
    theme_bw() +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme(legend.position = 'none',
          axis.title = element_blank(),
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 16, angle = 20, vjust = 0.9, hjust=0.75),
          panel.grid.minor = element_blank())
spp_vs_flamkm    
      
# Cluster plot          
km_clustflam_plot <- fviz_cluster(km, data = flamm_kmeans_scaled_df, geom = 'point', pointsize = 1.75, show.clust.cent = F, ellipse.type = 'norm') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
  scale_fill_manual(values = c(alpha("#00468BB2", 0.5), alpha('#ED0000B2', 0.5), alpha('#42B540B2', 0.5))) +
  theme_bw() +
  labs(color = 'K-Means Cluster', shape = 'K-Means Cluster', fill = 'K-Means Cluster') +
  theme(plot.title = element_blank(),
        legend.position = 'none',
        axis.title = element_text(face = 'bold', size = 17),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),  # Hide gridlines
        panel.grid.minor = element_blank())
km_clustflam_plot
```

### Arranging/Legend
```{r}
# For legend
km_leg_plot <- comp_kmeans_df %>% 
  ggplot(aes(x = cluster, y = lfm, color = as.factor(cluster))) +
  geom_point(shape = 'square', size = 5) +
  labs(color = 'K-Means Cluster') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
  theme_bw() +
  theme(legend.position = 'top',
        legend.title = element_text(face = 'bold', size = 25),
        legend.text = element_text(size = 20))
km_leg_plot
km_legend <- cowplot::get_legend(km_leg_plot)

mid <- cowplot::plot_grid(km_clustflam_plot, spp_vs_flamkm, rel_widths = c(1, 2.5), labels = c('A', 'B'), label_size = 24)
mid

arranged <- cowplot::plot_grid(km_legend, mid, km_boxplot_flam, rel_heights = c(1, 5, 5), ncol = 1, labels = c(' ', ' ', 'C'), label_size = 24)
arranged

#ggsave(here('figures', 'main_figures', 'flam_kmeans_cluster_arranged_figure.png'), height = 10, width = 18)
```

## Leaf Traits
```{r}
# labels
traits_labels <- c('Live Fuel Moisture (%)', 'Water Potential (-MPa)', 'Sample Weight (g)', 'Leaf Mass per Area', 'Leaf Thickness (mm)', 'Branching', 'Leaf Mass Ratio', 'Sample Density (g/cm3)', 'Sample Volume (cm3)')
names(traits_labels) <- c('lfm', 'mpa', 'sample_wt', 'LMA', 'thickness', 'branching', 'leaf_mass_ratio', 'sample_density', 'branch_volume')

# prepping dataset
traits_kmeans_df <- flamm.df %>% 
  filter(ignition != 0) %>% 
  select(species, species_id, ignition, lfm, mpa, sample_wt, LMA, thickness, branching, leaf_mass_ratio, sample_density, branch_volume) %>% 
  na.omit()

rownames(traits_kmeans_df) <- traits_kmeans_df$species_id

traits_kmeans_scaled_df <- traits_kmeans_df %>% 
  select(-c(species_id, species, ignition)) %>% 
  scale() %>% 
  as.data.frame()

# optimal number of clusters
# 2 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
fviz_nbclust(traits_kmeans_scaled_df, kmeans, method = "wss") #k = 3 or 4
# 2. silhouette method
fviz_nbclust(traits_kmeans_scaled_df, kmeans, method = "silhouette") #k = 4, 6, 8 or 10

# run k-means clustering
#perform k-means clustering with k = 3 clusters
km <- kmeans(traits_kmeans_scaled_df, centers = 3)
#plot results of final k-means model
fviz_cluster(km, data = traits_kmeans_scaled_df)
#add cluster assigment to original data
traits_kmeans_df <- cbind(traits_kmeans_df, cluster = km$cluster)

#visualizing how cluster assignment tracks species
traits_kmeans_df %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    scale_y_continuous(breaks = c(1:max(traits_kmeans_df$cluster))) +
    theme_bw() +
    labs(x = ' ', y = 'Cluster #', color = 'Cluster #', title = 'K-Means Clustering: Plant Traits')

traits_kmeans_df %>% 
  ggplot(aes(x = cluster, fill = species)) +
    geom_bar() +
    theme_bw() +
    labs(x = 'Cluster #', fill = 'Species', y = 'Count', title = 'K-Means Clustering: Plant Traits')

traits_kmeans_df %>% 
  pivot_longer(cols = c(lfm, mpa, sample_density, LMA, thickness, branching, leaf_mass_ratio, sample_wt, branch_volume), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(cluster), y = val, color = as.factor(cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free', labeller = labeller(var = traits_labels)) +
    labs(x = 'K-Means Cluster', y = 'Traits Value') +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 17),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))
```

## Traits and Flammability
```{r}
# labels
kmeans_labels <- c('Flame Height', 'Flame Duration', 'Temperature Change', 'Heat Flux Change', 'Live Fuel Moisture (%)', 'Water Potential (-MPa)', 'Sample Weight (g)', 'Leaf Mass per Area', 'Branching', 'Sample Density (g/cm3)')
names(kmeans_labels) <- c('fh', 'fd', 'temp_change', 'heat_flux_change', 'lfm', 'mpa', 'sample_wt', 'LMA', 'branching', 'sample_density')

# prepping dataset
kmeans_df <- flamm.df %>% 
  filter(ignition != 0) %>% 
  select(species, species_id, ignition, fh, fd, temp_change, heat_flux_change, lfm, mpa, sample_wt, LMA, branching, sample_density, species_id) %>% 
  na.omit()

rownames(kmeans_df) <- kmeans_df$species_id

kmeans_scaled_df <- kmeans_df %>% 
  select(-c(species_id, species, ignition, species_id)) %>% 
  scale() %>% 
  as.data.frame()

# optimal number of clusters
# 2 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
fviz_nbclust(kmeans_scaled_df, kmeans, method = "wss") #k = 2 or 3
# 2. silhouette method
fviz_nbclust(kmeans_scaled_df, kmeans, method = "silhouette") #k = 2

# run k-means clustering
#perform k-means clustering with k = 3 clusters
km <- kmeans(kmeans_scaled_df, centers = 2)

#plot results of final k-means model
fviz_cluster(km, data = kmeans_scaled_df) +
  scale_color_manual(values = c("#00468BB2", "#42B540B2")) +
  scale_fill_manual(values = c(alpha("#00468BB2", 0.5), alpha("#42B540B2", 0.5))) +
  theme_bw() +
  labs(color = 'K-Means Cluster', shape = 'K-Means Cluster', fill = 'K-Means Cluster') +
  theme(plot.title = element_blank(),
        legend.title = element_text(face = 'bold', size = 18),
        legend.text = element_text(size = 16),
        legend.position = c(0.15, 0.15),
        axis.title = element_text(face = 'bold', size = 17),
        axis.text = element_text(size = 15))
#ggsave(here('figures', 'main_figures', 'traits.flam.k.means.cluster.png'), width = 10, height = 7)

#add cluster assigment to original data
kmeans_df <- cbind(kmeans_df, cluster = km$cluster)

#visualizing how cluster assignment tracks species
kmeans_df %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    scale_y_continuous(breaks = c(1:max(traits_kmeans_df$cluster))) +
    labs(x = 'Species', y = 'K-Means Cluster', color = 'Cluster #') +
    theme_bw() +
    scale_color_manual(values = c("#00468BB2", "#42B540B2")) +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.text.y = element_text(size = 16), 
          axis.text.x = element_text(size = 14, angle = 20))
#ggsave(here('figures', 'main_figures', 'spp.vs.k.means.cluster.png'), height = 4.5, width = 9)

kmeans_df %>% 
  ggplot(aes(x = cluster, fill = species)) +
    geom_bar() +
    theme_bw() +
    labs(x = 'Cluster #', fill = 'Species', y = 'Count', title = 'K-Means Clustering')

plot.labels.df <- data.frame(var = c('branching', 'fd', 'fh', 'heat_flux_change', 'lfm', 'LMA', 'mpa', 'sample_density', 'sample_wt', 'temp_change'), label = c('A', 'B', 'C', 'D', 'E', 'F', 'G', 'H', 'I', 'J'))

#val = c(0.5, 95, 65, 0.23, 375, 0.265, 0.42, 0.235, 23.5, 38), cluster = c(rep(1, 10)),

kmeans_df %>% 
  filter(heat_flux_change < 0.3, sample_density < 0.4, temp_change < 50) %>% 
  pivot_longer(cols = c(fh, fd, temp_change, heat_flux_change, lfm, mpa, sample_wt, LMA, branching, sample_density), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(cluster), y = val, color = as.factor(cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free', labeller = labeller(var = kmeans_labels), ncol = 5) +
    geom_text(data = plot.labels.df, mapping = aes(x = -Inf, y = -Inf, label = label),
              color = 'black', hjust = -0.9, vjust = -11.2, size = 10, fontface = 'bold') +
    labs(x = 'K-Means Cluster') +
    scale_color_manual(values = c("#00468BB2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.x = element_text(face = 'bold', size = 17),
          axis.title.y = element_blank(),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))
#ggsave(here('figures', 'main_figures', 'traits.flam.k.means.boxplot.png'), width = 17, height = 9)

t.test(branching ~ cluster, data = kmeans_df) # sig. difference
t.test(lfm ~ cluster, data = kmeans_df) # sig. difference
t.test(mpa ~ cluster, data = kmeans_df) # sig. difference
t.test(LMA ~ cluster, data = kmeans_df) # sig. difference
t.test(sample_wt ~ cluster, data = kmeans_df) # sig. difference
t.test(sample_density ~ cluster, data = kmeans_df) # sig. difference
t.test(fh ~ cluster, data = kmeans_df) # sig. difference
t.test(fd ~ cluster, data = kmeans_df) # sig. difference
t.test(temp_change ~ cluster, data = kmeans_df) # sig. difference
t.test(heat_flux_change ~ cluster, data = kmeans_df) # sig. difference
```

### Intermediates Between Clusters
From above cluster plot, choosing samples that are in the 'intermediate space' between clusters (i.e., that area on the inner margins between the two clusters)
```{r}
kmeans_df <- kmeans_df %>% 
  mutate(int_cluster = ifelse(species_id %in% c('SALAPI_217', 'SALAPI_215', 'SALAPI_214', 'SALAPI_201', 'SALAPI_212', 'SALAPI_199', 'SALAPI_197', 'SALLEU_230', 'SALLEU_252', 'SALLEU_245', 'SALLEU_219', 'SALLEU_225', 'SALLEU_224', 'SALLEU_223', 'SALLEU_253', 'SALLEU_256', 'SALLEU_232', 'SALLEU_251', 'SALLEU_246', 'SALLEU_247', 'SALLEU_231', 'MALLAU_185', 'MALLAU_171', 'MALLAU_169', 'MALLAU_168', 'MALLAU_183', 'MALLAU_172', 'MALLAU_175', 'MALLAU_182', 'CEAGRI_86', 'CEAGRI_78', 'CEAGRI_72', 'ARCDEN_187', 'ARCDEN_188', 'ARCDEN_189', 'ARCDEN_190', 'ARCDEN_192', 'ARCDEN_193', 'ARCDEN_195', 'HETARB_138', 'HETARB_136'), 1.5, cluster)
  )

# species differences
kmeans_df %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = int_cluster, color = as.factor(int_cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    scale_y_continuous(breaks = c(1:3)) +
    labs(x = 'Species', y = 'K-Means Cluster', color = 'Cluster #') +
    theme_bw() +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.text.y = element_text(size = 16), 
          axis.text.x = element_text(size = 14, angle = 20))

kmeans_df %>% 
  filter(heat_flux_change < 0.3, sample_density < 0.4, temp_change < 50) %>% 
  pivot_longer(cols = c(fh, fd, temp_change, heat_flux_change, lfm, mpa, sample_wt, LMA, branching, sample_density), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(int_cluster), y = val, color = as.factor(int_cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free', labeller = labeller(var = kmeans_labels), ncol = 5) +
    geom_text(data = plot.labels.df, mapping = aes(x = -Inf, y = -Inf, label = label),
              color = 'black', hjust = -0.9, vjust = -11.2, size = 10, fontface = 'bold') +
    labs(x = 'K-Means Cluster') +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.x = element_text(face = 'bold', size = 17),
          axis.title.y = element_blank(),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))
```

## Traits and Flammability #2
```{r}
flamm.df %>% 
  ggplot(aes(x = branch_volume, y = sample_density)) +
    geom_point() +
    theme_bw()

flamm.df %>% 
  ggplot(aes(x = sample_wt, y = sample_density)) +
    geom_point() +
    theme_bw()

woohoo <- flamm.df %>% 
  select(species, sample_wt, branch_volume, sample_density, branch_length, branch_height, branch_width)
```


```{r}
set.seed(2)
# labels
kmeans_labels <- c('Flame Height', 'Flame Duration', 'Heat Flux Change', 'Live Fuel Moisture (%)', 'Dry Matter Content', 'Sample Weight (g)', 'Leaf Mass Ratio', 'Branching')
names(kmeans_labels) <- c('fh', 'fd', 'heat_flux_change', 'lfm', 'dmc', 'sample_wt', 'leaf_mass_ratio', 'branching')

# prepping dataset
kmeans_df2 <- flamm.df %>% 
  filter(ignition != 0) %>% 
  select(species, species_id, ignition, fh, fd, heat_flux_change, lfm, mpa, sample_wt, leaf_mass_ratio, branching, species_id) %>% 
  na.omit()

rownames(kmeans_df2) <- kmeans_df2$species_id

kmeans_scaled_df2 <- kmeans_df2 %>% 
  select(-c(species_id, species, ignition)) %>% 
  scale() %>% 
  as.data.frame()

# optimal number of clusters
# 2 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
fviz_nbclust(kmeans_scaled_df2, kmeans, method = "wss") #k = 3 or 4
# 2. silhouette method
fviz_nbclust(kmeans_scaled_df2, kmeans, method = "silhouette") #k = 2

# run k-means clustering
#perform k-means clustering with k = 2 clusters
km <- kmeans(kmeans_scaled_df2, centers = 3)
km$cluster <- case_when(km$cluster == 1 ~ 2,
                         km$cluster == 2 ~ 3,
                         km$cluster == 3 ~ 1)

#plot results of final k-means model
fviz_cluster(km, data = kmeans_scaled_df2) +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
  scale_fill_manual(values = c(alpha("#00468BB2", 0.5), alpha('#ED0000B2', 0.5), alpha('#42B540B2', 0.5))) +
  theme_bw() +
  labs(color = 'K-Means Cluster', shape = 'K-Means Cluster', fill = 'K-Means Cluster') +
  theme(plot.title = element_blank(),
        legend.title = element_text(face = 'bold', size = 18),
        legend.text = element_text(size = 16),
        legend.position = c(0.15, 0.15),
        axis.title = element_text(face = 'bold', size = 17),
        axis.text = element_text(size = 15))
#ggsave(here('figures', 'main_figures', 'traits.flam.k.means.cluster.no2.png'), width = 10, height = 7)

#add cluster assigment to original data
kmeans_df2 <- cbind(kmeans_df2, cluster = km$cluster)

# number of points per species, per cluster
number_pts_per_cluster <- kmeans_df2 %>% 
  dplyr::group_by(species, cluster) %>% 
  tally()

#visualizing how cluster assignment tracks species
kmeans_df2 %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    geom_text(aes(x = species, y = cluster + 0.25, color = as.factor(cluster), label = n),
                  data = number_pts_per_cluster, fontface = 'bold') +
    annotate('text', x = 0.55, y = 1.1, label = 'High Flam.', angle = 90, fontface = 'bold',
             size = 3.5, color = "#00468BB2") +
    annotate('text', x = 0.55, y = 2, label = 'Intermediate Flam.', angle = 90, fontface = 'bold',
             size = 3.5, color = "#ED0000B2") +
    annotate('text', x = 0.55, y = 3, label = 'Low Flam.', angle = 90, fontface = 'bold',
             size = 3.5, color = "#42B540B2") +
    scale_y_continuous(breaks = c(1:max(traits_kmeans_df$cluster))) +
    labs(x = 'Species', y = 'K-Means Cluster', color = 'Cluster #') +
    theme_bw() +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 18),
          axis.text.y = element_text(size = 16), 
          axis.text.x = element_text(size = 14, angle = 20, vjust = 0.9, hjust=0.75))
#ggsave(here('figures', 'main_figures', 'spp.vs.k.means.cluster.no2.png'), height = 4.5, width = 9)

plot.labels.df <- data.frame(var = c('branching', 'fd', 'fh', 'heat_flux_change', 'leaf_mass_ratio', 'lfm', 'mpa', 'sample_wt'), label = c('A', 'B', 'C', 'D', 'E', 'F', 'G', 'H'))

kmeans_df2 %>% 
  filter(heat_flux_change < 0.3) %>% 
  pivot_longer(cols = c(fh, fd, heat_flux_change, lfm, mpa, sample_wt, leaf_mass_ratio, branching), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = as.factor(cluster), y = val, color = as.factor(cluster))) +
    geom_boxplot(outlier.shape = NA, alpha= 0.8) +
    geom_jitter(width = 0.2, alpha = 0.2) +
    facet_wrap(~var, scales = 'free', labeller = labeller(var = kmeans_labels), ncol = 4) +
    geom_text(data = plot.labels.df, mapping = aes(x = -Inf, y = -Inf, label = label),
              color = 'black', hjust = -0.9, vjust = -11.2, size = 10, fontface = 'bold') +
    labs(x = 'K-Means Cluster') +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.x = element_text(face = 'bold', size = 17),
          axis.title.y = element_blank(),
          axis.text = element_text(size = 15),
          strip.text = element_text(face = 'bold', size = 15))
#ggsave(here('figures', 'main_figures', 'traits.flam.k.means.boxplot.no2.png'), width = 17, height = 9)
```

Alternative boxplot and reworking above figures to arrange:
```{r}
km_boxplotV2 <- kmeans_df2 %>% 
  mutate(fh = scale(fh), fd = scale(fd), heat_flux_change = scale(heat_flux_change),
         lfm = scale(lfm), mpa = scale(mpa), sample_wt = scale(sample_wt),
         leaf_mass_ratio = scale(leaf_mass_ratio), branching = scale(branching)) %>% 
  pivot_longer(cols = c(fh, fd, heat_flux_change, lfm, mpa, sample_wt, leaf_mass_ratio, branching),
               names_to = 'var', values_to = 'val') %>% 
  mutate(var = case_when(
    var == 'fh' ~ 'Flame Height',
    var == 'fd' ~ 'Flame Duration',
    var == 'heat_flux_change' ~ 'Heat Flux Change',
    var == 'lfm' ~ 'Live Fuel Moisture',
    var == 'mpa' ~ 'Water Potential',
    var == 'sample_wt' ~ 'Sample Weight',
    var == 'leaf_mass_ratio' ~ 'Leaf Mass Ratio',
    var == 'branching' ~ 'Branching')) %>% 
  ggplot(aes(x = var, y = val, color = as.factor(cluster))) +
    geom_boxplot(alpha= 0.8) +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.text.x = element_text(face = 'bold', size = 15),
          axis.title = element_blank(),
          axis.text.y = element_text(size = 15))
km_boxplotV2
#ggsave(here('figures', 'main_figures', 'traits.flam.k.means.boxplotV2.no2.png'), width = 16, height = 8)

spp_vs_km <- kmeans_df2 %>% 
  mutate(species = fct_relevel(species, c('HETARB', 'MALLAU', 'CEAGRI', 'SALAPI', 'SALLEU',
                                                 'ARCDEN', 'ARTCAL', 'ERIKAR'))) %>% 
  ggplot(aes(x = species, y = cluster, color = as.factor(cluster))) +
    geom_jitter(width = 0.1, height = 0.1) +
    geom_text(aes(x = species, y = cluster + 0.25, color = as.factor(cluster), label = n),
                  data = number_pts_per_cluster, fontface = 'bold', size = 5.5) +
    annotate('text', x = 0.55, y = 1, label = 'High', angle = 90, fontface = 'bold',
             size = 5.5, color = "#00468BB2") +
    annotate('text', x = 0.8, y = 2, label = 'Flammability', angle = 90, fontface = 'bold',
             size = 5.5, color = "black") +
    annotate('text', x = 0.55, y = 3, label = 'Low', angle = 90, fontface = 'bold',
             size = 5.5, color = "#42B540B2") +
    geom_segment(aes(x = 0.55, y = 1.6, xend = 0.55, yend = 1.2),
                  arrow = arrow(length = unit(0.2, "cm")), color = "#00468BB2") +
    geom_segment(aes(x = 0.55, y = 1.6, xend = 0.55, yend = 2.4), color = "#ED0000B2") +
    geom_segment(aes(x = 0.55, y = 2.4, xend = 0.55, yend = 2.8),
                  arrow = arrow(length = unit(0.2, "cm")), color = "#42B540B2") +
    scale_y_continuous(breaks = c(1:max(traits_kmeans_df$cluster))) +
    theme_bw() +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
    theme(legend.position = 'none',
          axis.title = element_blank(),
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 16, angle = 20, vjust = 0.9, hjust=0.75),
          panel.grid.minor = element_blank())
spp_vs_km   
                
km_clust_plot <- fviz_cluster(km, data = kmeans_scaled_df2, geom = 'point', pointsize = 1.75, show.clust.cent = F, ellipse.type = 'norm') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
  scale_fill_manual(values = c(alpha("#00468BB2", 0.5), alpha('#ED0000B2', 0.5), alpha('#42B540B2', 0.5))) +
  theme_bw() +
  labs(x = 'Dimension 1 (42.5%)', y = 'Dimension 2 (17.9%)', color = 'K-Means Cluster', shape = 'K-Means Cluster', fill = 'K-Means Cluster') +
  theme(plot.title = element_blank(),
        legend.position = 'none',
        axis.title = element_text(face = 'bold', size = 17),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),  # Hide gridlines
        panel.grid.minor = element_blank())
km_clust_plot
```

### Arranging/Legend
```{r}
# For legend
km_leg_plot <- kmeans_df2 %>% 
  ggplot(aes(x = cluster, y = lfm, color = as.factor(cluster))) +
  geom_point(shape = 'square', size = 5) +
  labs(color = 'K-Means Cluster') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2")) +
  theme_bw() +
  theme(legend.position = 'top',
        legend.title = element_text(face = 'bold', size = 25),
        legend.text = element_text(size = 20))
km_leg_plot
km_legend <- cowplot::get_legend(km_leg_plot)

mid <- cowplot::plot_grid(km_clust_plot, spp_vs_km, rel_widths = c(1, 2.5), labels = c('A', 'B'), label_size = 24)
mid

arranged <- cowplot::plot_grid(km_legend, mid, km_boxplotV2, rel_heights = c(1, 5, 5), ncol = 1, labels = c(' ', ' ', 'C'), label_size = 24)
arranged

ggsave(here('figures', 'supp_figures', 'kmeans_cluster_interspecific.png'), height = 10, width = 16)
```
