---
title: "Principal Component Analysis"
author: "Joe Celebrezze"
date: "2023-07-05"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
here = here::here
source(here("scripts", "source_code.R"))
```

# Data Wrangling 
```{r}
flamm.df <- flamm.df %>% 
  filter(species %in% c('ARTCAL', 'CEAGRI', 'ERIKAR', 'MALLAU', 'SALAPI', 'SALLEU', 'ARCDEN', 'HETARB')) %>% 
  filter(ignition != '0') %>% 
  filter(temp_change < 100) %>%  # removing outlier, as it makes things funky
  unite('id', c(species, plant, rep), sep = '_', remove = F)
```


# 1. Flammability Metrics
## 1.1. PCA Figure
Plot setup:
```{r warning=FALSE}
flamm.df_pca <- flamm.df %>% 
  filter(ignition != 0) %>% # No non-ignitions
   select(fh, fd, temp_change, heat_flux_change, species, id) %>% 
   drop_na(fh, fd, temp_change, heat_flux_change, species)

figure_pca_quant <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species")]) %>%  
 transmute("Flame Duration" = fd, #using spaces to shift labels over instead of \t since \t is leading to little blank squares popping up on the plot
           "Flame Height" = fh,
           "Temp Change" = temp_change,
           "Heat Flux Change" = heat_flux_change
          ) %>% 
  na.omit()

figure_pca_cat <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species")]) %>%  
  na.omit() %>% 
  select(species) %>% 
  transmute("Species" = species)
  
figure_prcomp <- prcomp(na.omit(figure_pca_quant), 
                          center = TRUE, 
                          scale = TRUE, 
                          retx=TRUE)
  
figure_pca <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species")]) %>%  
  na.omit() 
```

```{r warning=FALSE}
#Setting up plotting data: 

#decisions for plotting:
choices = 1:2 
scale = 1
obs.scale = 1 - scale
var.scale = scale
ellipse.prob = 0.68
labels.size = 3
circle.prob = 0.69
choices = 1:2

#Run PCA
pcobj <- prcomp(na.omit(figure_pca_quant), center = TRUE, scale = TRUE)

# extract PCA components: 
nobs.factor <- sqrt(nrow(pcobj$x) - 1)
    d <- pcobj$sdev
    u <- sweep(pcobj$x, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$rotation

  choices <- pmin(choices, ncol(u))
  df.u <- as.data.frame(sweep(u[, choices], 2, d[choices]^obs.scale, 
                              FUN = "*"))
  v <- sweep(v, 2, d^var.scale, FUN = "*")
  df.v <- as.data.frame(v[, choices])
  names(df.u) <- c("PC1", "PC2")
  names(df.v) <- names(df.u)
  df.u <- df.u * nobs.factor

  r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(df.u^2))^(1/4)
  
v.scale <- rowSums(v^2)
df.v <- r * df.v/sqrt(max(v.scale)) 
df.v <- df.v %>% mutate(PC1, PC2)
df.v$Variables <- rownames(v) 
PCAloadings = df.v
#df.v = dataset with loadings 

#dataset with scores and categorical variables for plotting points: 
PCAvalues <- cbind(df.u, figure_pca_cat)

#dataset with information for plotting circle: 
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))

circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta)) 

#Group by flam. metric: 
PCA_metrics <- df.v %>% 
  filter(Variables %in% c("Flame Duration", "Flame Height", "Temp Change",
                          "Heat Flux Change"))
    
# Calculate the angles and the label offset
df.v$Angle = ((180/pi) * atan(df.v$PC2/df.v$PC1))

df.v$Offset <- ((-2 * sign(df.v$PC1))/2)

#labels: 
u.axis.labs <- paste("PC", choices, sep = "")
u.axis.labs <- paste(u.axis.labs, sprintf("(%0.1f%% explained var.)", 
                                            100 * pcobj$sdev[choices]^2/sum(pcobj$sdev^2)))
```

Plot:
```{r}
flam.pca <- ggplot(data = PCAvalues, aes(x = PC1, y = PC2)) +
  ylim(-3.5, 3.5) +
  geom_point(size = 1.75, alpha = 0.6, aes(color = Species), data = PCAvalues) +
  stat_ellipse(aes(color = Species), alpha = 0.6) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  geom_segment(data = PCA_metrics, aes(x = 0, y = 0, 
                                              xend = PC1, yend = PC2),
     arrow = arrow(length = unit(1/2, "picas")), 
    linewidth = .9, alpha = 0.9,
     color = "black") +
  annotate("text", 
          x = (PCA_metrics$PC1 * 1.65), 
          y = (PCA_metrics$PC2 * 1.2),
          label = PCA_metrics$Variables, size = 6,
          fontface = 'bold') +
  theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
          panel.grid.major = element_blank(),  # Hide major gridlines
          panel.grid.minor = element_blank(),
          legend.key = element_rect(fill = "white"),
          legend.position = 'none', # change to 'right' if you want figure alone, change to 'none' if you want to arrange figures
          legend.title = element_text(face = 'bold', size = 15),
          legend.text = element_text(face = 'italic', size = 13),
          axis.title = element_text(face = 'bold', size = 20),
          axis.text = element_text(size = 18),
          plot.margin = unit(c(.15, .15, .225, .3), "cm")) +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
  geom_path(data = circle, color = "black",  size = 1/2, alpha = 1/3) + 
  xlab(u.axis.labs[1]) + 
   ylab(u.axis.labs[2]) #+
  #coord_equal() #remove hash if you want to make coordinates equal; for arranging figure, looks best if they're not
flam.pca
#ggsave(here('figures', 'PCA', 'Flamm.PCA.png'), height = 7, width = 9)
```

## 1.2. PCA Table

```{r warning=FALSE}
figure_varimax <- varimax(pcobj$rotation[,1:3])
figure_varimax

tab_pca(pcobj, rotation = 'varimax',
        nmbr.fctr = 4,
        digits = 2,
        show.var = TRUE,
  string.pov = "Proportion of Variance",
  string.cpov = "Cumulative Proportion",
  file = here('figures', 'PCA', 'Flamm.PC.Table.html'))
```

## Adding PC Scores to Data
```{r warning=FALSE}
x_flam <-predict(pcobj)  ### generate scores for each row in flamm.df

pca_flamm_data <- as_data_frame(x_flam) 

pca_df <- cbind(pca_flamm_data, flamm.df_pca)
```

# 2. Plant Traits
## 2.1. PCA Figure
Plot setup:
```{r warning=FALSE}
traits.df_pca <- flamm.df %>% 
  filter(ignition != 0) %>% # No non-ignitions
   select(id, species, lfm, mpa, sample_density, LMA, thickness, branching, leaf_mass_ratio, sample_wt, leaf_sav, dmc, stem_sav) %>% 
   drop_na(species, lfm, mpa, sample_density, LMA, thickness, branching, leaf_mass_ratio, sample_wt, leaf_sav, dmc, stem_sav) %>% 
  select(id, species, lfm, mpa, LMA, thickness, branching, leaf_mass_ratio, leaf_sav, dmc, stem_sav)

figure_pca_quant2 <- (traits.df_pca[, c('lfm', 'mpa', 'LMA', 'thickness', 'branching', 'leaf_mass_ratio', 'leaf_sav', 'dmc', 'stem_sav')]) %>%  
 transmute("LFM" = lfm,  # added spaces to adjust label position
           "Water Potential" = mpa,
           "LMA" = LMA,
           "     Thickness" = thickness,
           "Branching" = branching,
           "Leaf Ratio" = leaf_mass_ratio,
           #"Sample Weight" = sample_wt,
           "\nDMC     " = dmc, # added spaces to adjust label position
           "\nLeaf SA:V" = leaf_sav,
           "Stem SA:V" = stem_sav) %>% 
  na.omit()

figure_pca_cat2 <- (traits.df_pca[, c('species', 'lfm', 'mpa', 'LMA', 'thickness', 'branching', 'leaf_mass_ratio', 'leaf_sav', 'dmc', 'stem_sav')]) %>% 
  na.omit() %>% 
  select(species) %>% 
  transmute("Species" = species)
  
figure_prcomp2 <- prcomp(na.omit(figure_pca_quant2), 
                          center = TRUE, 
                          scale = TRUE, 
                          retx=TRUE)
  
figure_pca2 <- (traits.df_pca[, c('species', 'lfm', 'mpa', 'LMA', 'thickness', 'branching', 'leaf_mass_ratio', 'leaf_sav', 'dmc', 'stem_sav')]) %>%  
  na.omit() 
```

```{r warning=FALSE}
#Setting up plotting data: 

#decisions for plotting:
choices = 1:2 
scale = 1
obs.scale = 1 - scale
var.scale = scale
ellipse.prob = 0.68
labels.size = 3
circle.prob = 0.69
choices = 1:2

#Run PCA
pcobj2 <- prcomp(na.omit(figure_pca_quant2), center = TRUE, scale = TRUE)

# extract PCA components: 
nobs.factor <- sqrt(nrow(pcobj2$x) - 1)
    d <- pcobj2$sdev
    u <- sweep(pcobj2$x, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj2$rotation

  choices <- pmin(choices, ncol(u))
  df.u <- as.data.frame(sweep(u[, choices], 2, d[choices]^obs.scale, 
                              FUN = "*"))
  v <- sweep(v, 2, d^var.scale, FUN = "*")
  df.v <- as.data.frame(v[, choices])
  names(df.u) <- c("PC1", "PC2")
  names(df.v) <- names(df.u)
  df.u <- df.u * nobs.factor

  r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(df.u^2))^(1/4)
  
v.scale <- rowSums(v^2)
df.v <- r * df.v/sqrt(max(v.scale)) 
df.v <- df.v %>% mutate(PC1, PC2)
df.v$Variables <- rownames(v) 
PCAloadings = df.v
#df.v = dataset with loadings 

#dataset with scores and categorical variables for plotting points: 
PCAvalues2 <- cbind(df.u, figure_pca_cat2)

#dataset with information for plotting circle: 
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))

circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta)) 

#Group by flam. metric: 
PCA_metrics2 <- df.v %>% 
  filter(Variables %in% c("LFM", "Water Potential", "Sample Density", "LMA",
           "     Thickness", "Branching", "Leaf Ratio", "\nDMC     ",
           "\nLeaf SA:V", "Stem SA:V"))
    
# Calculate the angles and the label offset
df.v$Angle = ((180/pi) * atan(df.v$PC2/df.v$PC1))

df.v$Offset <- ((-2 * sign(df.v$PC1))/2)

#labels: 
u.axis.labs <- paste("PC", choices, sep = "")
u.axis.labs <- paste(u.axis.labs, sprintf("(%0.1f%% explained var.)", 
                                            100 * pcobj2$sdev[choices]^2/sum(pcobj2$sdev^2)))
```

Plot:
```{r}
traits.pca <- ggplot(PCAvalues2, aes(x = PC1, y = PC2)) +
  ylim(-3.8, 2.4) +
  xlim(-2.1, 2.4) +
  geom_point(size = 1.75, alpha = 0.6, aes(color = Species), data = PCAvalues2) +
  stat_ellipse(aes(color = Species), alpha = 0.6) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  geom_segment(data = PCA_metrics2, aes(x = 0, y = 0, 
                                              xend = PC1, yend = PC2),
     arrow = arrow(length = unit(1/2, "picas")), 
    linewidth = .9, alpha = 0.9,
     color = "black") +
  annotate("text", 
          x = (PCA_metrics2$PC1 * 1.34), 
          y = (PCA_metrics2$PC2 * 1.2),
          label = PCA_metrics2$Variables, size = 6,
          fontface = 'bold') +
  theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
          panel.grid.major = element_blank(),  # Hide major gridlines
          panel.grid.minor = element_blank(),
          legend.key = element_rect(fill = "white"),
          legend.position = 'none', # change to 'right' if you want figure alone, change to 'none' if you want to arrange figures
          legend.title = element_text(face = 'bold', size = 15),
          legend.text = element_text(face = 'italic', size = 13),
          axis.title = element_text(face = 'bold', size = 20),
          axis.text = element_text(size = 18),
          plot.margin = unit(c(.25, .15, .15, .2), "cm")) + 
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
  xlab(u.axis.labs[1]) + 
  ylab(u.axis.labs[2])# +
  #coord_equal() #remove hash if you want to make coordinates equal; for arranging figure, looks best if they're not
traits.pca
#ggsave(here('figures', 'PCA', 'Traits.PCA.png'), height = 7, width = 9)
```

## 2.2. PCA Table
```{r warning=FALSE}
figure_varimax <- varimax(pcobj2$rotation[,1:3])
figure_varimax

summary(pcobj2)

tab_pca(pcobj2, rotation = 'varimax',
        nmbr.fctr = 4,
        digits = 2,
        show.var = TRUE,
  string.pov = "Proportion of Variance",
  string.cpov = "Cumulative Proportion",
  file = here('figures', 'PCA', 'Traits.PC.Table.html'))
```

## Adding PC Scores to Data
```{r warning=FALSE}
x_traits <-predict(pcobj2)  ### generate scores for each row in flamm.df

pca_traits_data <- as_data_frame(x_traits) 

pca_df2 <- cbind(pca_traits_data, traits.df_pca)
```

# 3. Comparing Principal Components
```{r}
pca_df <- pca_df %>% 
  #select(-PC4) %>%  # Not interested in PC4 or PC5 (>80% of variance explained by first three PCs)
  rename(PC1_flamm = PC1,
         PC2_flamm = PC2,
         PC3_flamm = PC3,
         PC4_flamm = PC4)

pca_df2 <- pca_df2 %>% 
  select(-PC5, -PC6, -PC7, -PC8, -PC9, -species) %>%  # Not interested in PC5 through PC10 (>80% of variance explained by first four PCs); species variable already in pca_df
  rename(PC1_traits = PC1,
         PC2_traits = PC2,
         PC3_traits = PC3,
         PC4_traits = PC4)

pca_df_main <- merge(pca_df, pca_df2, by = 'id')
```

## Scatterplots
### Stand-alone
```{r}
pc_labels <- c('PC1 (-FH)', 'PC2 (-\U0394Temp)', 'PC3 (FD)', 'PC4 (-\U0394Heat Flux)', 'PC1 (DMC, -Density, -LMR)', 'PC2 (Leaf SA:V, -Thickness)', 'PC3 (LMA, Sample Weight)', 'PC4 (-LFM, MPa, Branching)')
names(pc_labels) <- c('PC1_flamm', 'PC2_flamm', 'PC3_flamm', 'PC4_flamm', 'PC1_traits', 'PC2_traits', 'PC3_traits', 'PC4_traits')

pca_df_main %>% 
  pivot_longer(cols = c(PC1_flamm, PC2_flamm, PC3_flamm, PC4_flamm), values_to = 'flamm_pc_score', names_to = 'flamm_pc') %>% 
  pivot_longer(cols = c(PC1_traits, PC2_traits, PC3_traits, PC4_traits), values_to = 'traits_pc_score', names_to = 'traits_pc') %>% 
  ggplot(aes(x = traits_pc_score, y = flamm_pc_score)) +
    geom_point(alpha = 0.5, aes(color = species)) +
    geom_smooth(method = 'lm', color = 'black', alpha = 0.75) +
    stat_cor(label.y.npc = 'bottom', label.x.npc = 'left', size = 2.5)+ 
    #stat_regline_equation(label.y.npc = 'bottom', label.x.npc = 'left', size = 3) +
    facet_grid(flamm_pc ~ traits_pc, scales = 'free', labeller = labeller(flamm_pc = pc_labels, traits_pc = pc_labels)) +
    labs(y = 'Flamm. PC Score', x = 'Traits PC Score', color = 'Species') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
    theme_bw() +
    theme(legend.position = 'right',
          legend.title = element_text(size = 16, face = 'bold'),
          axis.title = element_text(size = 16, face = 'bold'),
          axis.text = element_text(size = 13),
          strip.text = element_text(size = 11.5, face = 'bold'))
ggsave(here('figures', 'supp_figures', 'PC.Regressions.png'), height = 8, width = 12)
```

### For Arranged Main Figure
3x3 panels
```{r}
pc_labels <- c('PC1 (-FH)', 'PC2 (-\U0394Temp)', 'PC3 (FD)', 'PC1 (-LFM, -LMR, DMC)', 'PC2 (-MPa, -Branching)', 'PC3 (-LTh, Leaf SA:V)')
names(pc_labels) <- c('PC1_flamm', 'PC2_flamm', 'PC3_flamm', 'PC1_traits', 'PC2_traits', 'PC3_traits')

pc.scatterplot <- pca_df_main %>% 
  pivot_longer(cols = c(PC1_flamm, PC2_flamm, PC3_flamm), values_to = 'flamm_pc_score', names_to = 'flamm_pc') %>% 
  pivot_longer(cols = c(PC1_traits, PC2_traits, PC3_traits), values_to = 'traits_pc_score', names_to = 'traits_pc') %>% 
  ggplot(aes(x = traits_pc_score, y = flamm_pc_score)) +
    geom_point(alpha = 0.5, aes(color = species)) +
    geom_smooth(method = 'lm', color = 'black', alpha = 0.75) +
    stat_cor(label.y.npc = 'bottom', label.x.npc = 'left', size = 5, fontface = 'bold')+ 
    #stat_regline_equation(label.y.npc = 'bottom', label.x.npc = 'left', size = 3) +
    facet_grid(flamm_pc ~ traits_pc, scales = 'free', labeller = labeller(flamm_pc = pc_labels, traits_pc = pc_labels)) +
    labs(y = 'Flamm. PC Score', x = 'Traits PC Score') +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(size = 18, face = 'bold'),
          axis.text = element_text(size = 14),
          strip.text = element_text(size = 14, face = 'bold'))
pc.scatterplot
```

PC1 vs PC1
```{r}
pc1.scatterplot <- pca_df_main %>% 
  ggplot(aes(x = PC1_traits, y = PC1_flamm)) +
    geom_point(alpha = 0.6, aes(color = species)) +
    geom_smooth(method = 'lm', color = 'black', alpha = 0.75, fill = 'gray75') +
    stat_cor(label.y.npc = 'bottom', label.x.npc = 'left', size = 6, fontface = 'bold')+ 
    #stat_regline_equation(label.y.npc = 'bottom', label.x.npc = 'left', size = 3) +
    labs(y = 'Flammability PC1', x = 'Traits PC1') +
    scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2",
                                  "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
    scale_y_continuous(breaks = c(-4, -2, 0, 2)) +
    scale_x_continuous(breaks = c(-2, 0, 2, 4)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(size = 20, face = 'bold'),
          axis.text = element_text(size = 18))
pc1.scatterplot
```

## Second-order PCA
```{r}
# selecting metrics of interest
second_order_pca_df <- pca_df_main %>% 
  select(species, PC1_flamm, PC2_flamm, PC3_flamm, PC1_traits, PC2_traits, PC3_traits, PC4_traits) # selecting principal components from above

# looking at correlation and distribution
pairs.panels(second_order_pca_df)

# prepping dataset for PCA
second_order_pca_df <- second_order_pca_df %>% 
  na.omit()

# running PCA
second_order_pcobj <- second_order_pca_df %>% 
  select(-species) %>% 
  prcomp(center = TRUE, scale = TRUE) 

# first glance at results
summary(second_order_pcobj)
plot(second_order_pcobj, type = 'lines') # first 2 explain 51.86% variance
print(second_order_pcobj) # PC1: flame height, flame duration, max temp and max heat flux (combustibility), PC2: time to ignition, proportion ignited, post-flame glow (ignitability and consumability)

# visualizing results
ggbiplot(second_order_pcobj,
         groups = second_order_pca_df$species,
         ellipse = TRUE, varname.size = 5, alpha = 0.5) +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
    theme_bw()

# eigenvalue table
tab_pca(second_order_pcobj, rotation = 'varimax',
        nmbr.fctr = 5,
        digits = 2,
        show.var = TRUE,
  string.pov = "Proportion of Variance",
  string.cpov = "Cumulative Proportion") 
```

# 4. Main Figure: Arranging Figures
## Legend
```{r}
PCAvalues_legend <- PCAvalues2 %>% 
  mutate(Species = case_when(Species == 'ARCDEN' ~ 'Ar. densiflora',
                             Species == 'CEAGRI' ~ 'Ce. griseus',
                             Species == 'ARTCAL' ~ 'Ar. californica',
                             Species == 'ERIKAR' ~ 'Er. karvinskianus',
                             Species == 'HETARB' ~ 'He. arbutifolia',
                             Species == 'MALLAU' ~ 'Ma. laurina',
                             Species == 'SALAPI' ~ 'Sa. apiana',
                             Species == 'SALLEU' ~ 'Sa. leucophylla'))

legend <- get_legend(ggplot(PCAvalues_legend, aes(x = PC1, y = PC2)) +
  geom_point(size = 1.75, alpha = .9, aes(color = Species), data = PCAvalues_legend) +
  stat_ellipse(aes(color = Species)) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme(legend.key = element_rect(fill = "white"),
        legend.position = 'bottom',
        legend.title = element_text(face = 'bold', size = 18),
        legend.text = element_text(face = 'italic', size = 16)) +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")))
```

## Arranged
```{r}
# blank figure
blank <- ggplot(data = flamm.df, aes(x = species, y = mpa)) +
  theme_void()

# arranging
plots <- ggarrange(traits.pca, blank, pc.scatterplot, flam.pca,
          labels = c('A', ' ', 'C', 'B'),
          font.label = list(size = 34),
          ncol = 2, nrow = 2,
          heights = c(2, 2), widths = c(5, 4))
ggarrange(plots, legend,
          ncol = 1, heights = c(8, 1))
ggsave(here('figures', 'extra_figures', 'pca.and.scatterplots.png'), height = 14, width = 16)

plots <- ggarrange(flam.pca, pc1.scatterplot, blank, traits.pca,
          labels = c('a', 'c', '', 'b'),
          font.label = list(size = 40),
          ncol = 2, nrow = 2,
          heights = c(2, 2), widths = c(5, 4))
ggarrange(plots, legend,
          ncol = 1, heights = c(8, 1))
ggsave(here('figures', 'main_figures', 'pc1.vs.pc1.and.scatterplots.png'), height = 14, width = 16)
```


# ----------------------------------------
# 5. Flam. and Plant Traits PCA
Using key variables identified in above PCAs and in MEM selection 

## 5.1. PCA Figure
Plot setup:
```{r warning=FALSE}
flamm.df_pca <- flamm.df %>% 
  filter(ignition != 0) %>% # No non-ignitions
   select(fh, fd, temp_change, heat_flux_change, lfm, mpa, sample_wt, sample_density, LMA, branching, species, id) %>% 
   drop_na(fh, fd, temp_change, heat_flux_change, lfm, mpa, sample_wt, sample_density, LMA, branching, species)

figure_pca_quant <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species",'lfm', 'mpa', 'sample_wt', 'sample_density', 'LMA', 'branching')]) %>%  
 transmute("Flame Duration" = fd, #using spaces to shift labels over instead of \t since \t is leading to little blank squares popping up on the plot
           "Flame Height" = fh,
           "Temp Change" = temp_change,
           "Heat Flux Change" = heat_flux_change,
           "Live Fuel Moisture" = lfm, #using spaces to shift labels over instead of \t since \t is leading to little blank squares popping up on the plot
           "Water Potential" = mpa,
           "Sample Weight (g)" = sample_wt,
           "Sample Density (g/cm3)" = sample_density,
           "Leaf Mass per Area" = LMA,
           "Branching" = branching,
          ) %>% 
  na.omit()

figure_pca_cat <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species",'lfm', 'mpa', 'sample_wt', 'sample_density', 'LMA', 'branching')]) %>%  
  na.omit() %>% 
  select(species) %>% 
  transmute("Species" = species)
  
figure_prcomp <- prcomp(na.omit(figure_pca_quant), 
                          center = TRUE, 
                          scale = TRUE, 
                          retx=TRUE)
  
figure_pca <- (flamm.df_pca[, c("fh", "fd", "temp_change", "heat_flux_change", "species",'lfm', 'mpa', 'sample_wt', 'sample_density', 'LMA', 'branching')]) %>%  
  na.omit() 
```

```{r warning=FALSE}
#Setting up plotting data: 

#decisions for plotting:
choices = 1:2 
scale = 1
obs.scale = 1 - scale
var.scale = scale
ellipse.prob = 0.68
labels.size = 3
circle.prob = 0.69
choices = 1:2

#Run PCA
pcobj <- prcomp(na.omit(figure_pca_quant), center = TRUE, scale = TRUE)

# extract PCA components: 
nobs.factor <- sqrt(nrow(pcobj$x) - 1)
    d <- pcobj$sdev
    u <- sweep(pcobj$x, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$rotation

  choices <- pmin(choices, ncol(u))
  df.u <- as.data.frame(sweep(u[, choices], 2, d[choices]^obs.scale, 
                              FUN = "*"))
  v <- sweep(v, 2, d^var.scale, FUN = "*")
  df.v <- as.data.frame(v[, choices])
  names(df.u) <- c("PC1", "PC2")
  names(df.v) <- names(df.u)
  df.u <- df.u * nobs.factor

  r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(df.u^2))^(1/4)
  
v.scale <- rowSums(v^2)
df.v <- r * df.v/sqrt(max(v.scale)) 
df.v <- df.v %>% mutate(PC1, PC2)
df.v$Variables <- rownames(v) 
PCAloadings = df.v
#df.v = dataset with loadings 

#dataset with scores and categorical variables for plotting points: 
PCAvalues <- cbind(df.u, figure_pca_cat)

#dataset with information for plotting circle: 
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))

circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta)) 

#Group by flam. metric: 
PCA_metrics <- df.v %>% 
  filter(Variables %in% c("Flame Duration", "Flame Height", "Temp Change",
                          "Heat Flux Change", "Live Fuel Moisture",
                          "Water Potential", "Sample Weight (g)",
                          "Sample Weight (g/cm3)",
                          "Leaf Mass per Area", "Branching"))
    
# Calculate the angles and the label offset
df.v$Angle = ((180/pi) * atan(df.v$PC2/df.v$PC1))

df.v$Offset <- ((-2 * sign(df.v$PC1))/2)

#labels: 
u.axis.labs <- paste("PC", choices, sep = "")
u.axis.labs <- paste(u.axis.labs, sprintf("(%0.1f%% explained var.)", 
                                            100 * pcobj$sdev[choices]^2/sum(pcobj$sdev^2)))
```

Plot:
```{r}
ggplot(PCAvalues, aes(x = PC1, y = PC2)) +
  ylim(-4, 3) +
  geom_point(size = 1.75, alpha = .9, aes(color = Species), data = PCAvalues) +
  stat_ellipse(aes(color = Species)) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  geom_segment(data = PCA_metrics, aes(x = 0, y = 0, 
                                              xend = PC1, yend = PC2),
     arrow = arrow(length = unit(1/2, "picas")), 
    linewidth = .9, alpha = 0.9,
     color = "black") +
  annotate("text", 
          x = (PCA_metrics$PC1 * 1.3), 
          y = (PCA_metrics$PC2 * 1.1),
          label = PCA_metrics$Variables, size = 3.5,
          fontface = 'bold') +
  theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
          panel.grid.major = element_blank(),  # Hide major gridlines
          panel.grid.minor = element_blank(),
          legend.key = element_rect(fill = "white"),
          legend.position = 'right',
          legend.title = element_text(face = 'bold', size = 15),
          legend.text = element_text(face = 'italic', size = 13),
          axis.title = element_text(face = 'bold', size = 15)) +
  scale_color_manual(values = c("#00468BB2", "#ED0000B2", "#42B540B2", "#0099B4B2", "#925E9FB2", "#AD002AB2", "#ADB6B6B2", "#1B1919B2")) +
   geom_path(data = circle, color = "black",  size = 1/2, alpha = 1/3) + 
   xlab(u.axis.labs[1]) + 
   ylab(u.axis.labs[2]) +
  coord_equal()
#ggsave(here('figures', 'PCA', 'FlammandTraits.PCA.png'), height = 7, width = 9)
```

## 5.2. PCA Table
```{r warning=FALSE}
figure_varimax <- varimax(pcobj2$rotation[,1:3])
figure_varimax

#tab_pca(pcobj, rotation = 'varimax',
#        nmbr.fctr = 4,
#         digits = 2,
#         show.var = TRUE,
#  string.pov = "Proportion of Variance",
#  string.cpov = "Cumulative Proportion",
#  file = here('figures', 'PCA', 'FlammandTraits.PC.Table.html')) 
```

