---
title: "Data Wrangling: Plant Traits"
author: "Zoe Fung & Joe Celebrezze"
date: "2023-04-17"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(tidyverse)
library(magrittr)
library(janitor)
library(dplyr)
library(here)
library(bayestestR) # For area_under_curve

# Functions
rename = dplyr::rename
here = here::here
summarise = dplyr::summarise
group_by = dplyr::group_by
select = dplyr::select
```

# Initial Data Cleaning
## Leaf Data
```{r}
# load in data
leaf_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_data_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(leaf_data)
```

## Leaf Area
```{r}
# load in data
leaf_area <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_area_flamm.csv')) %>% 
  clean_names() %>% 
  select(-x1, -x_1)

# view column names and data types
glimpse(leaf_area)
```

## Stem-Leaf Ratio
```{r}
# load in data
stem_leaf_ratio <- read_csv(here('data', 'raw-data', 'plant-traits', 'stemleaf_massratio_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(stem_leaf_ratio)
```

## Stem Data
```{r}
# load in data
stem_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'stem_data_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(stem_data)
```

## VOC Data
```{r}
voc_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'VOCs_JL.csv')) %>%
  clean_names() %>% 
  filter(species != 'Arctostaphylos densifolia', # Not interested in species we didn't burn
         method %in% c('Control', 'Heated'), # Only interested in these two methods
         treatment == 'WET')  %>%  # Only interested in wet treatment (since we didn't do droughted for the flammability testing)
  select(-treatment) # Since we discarded droughted treatment, we don't really need that column anymore
```


# Calculating LFM
First, getting cup weights data
```{r}
cup_weights_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'cup_weights.csv')) %>% 
  slice_head(n = 82)  %>% 
  rename(cup_mass = mass) %>% 
  mutate(dry_cup_mass = (mass_post_drying + mass_post_drying2)/2) %>%  # Using average dry cup mass as 'dry cup mass'
  select(-mass_post_drying, -mass_post_drying2)

leaf_data <- merge(leaf_data, cup_weights_data, by = 'cup_number')
stem_data <- merge(stem_data, cup_weights_data, by = 'cup_number')
```

Subtracting cup weight from fresh and dry weights and then calculating stem- and leaf-specific LFM
```{r}
leaf_data <- leaf_data %>% 
  mutate(leaf_wet_mass_g = ifelse(leaf_wet_mass_g > 1000, leaf_wet_mass_g/1000, leaf_wet_mass_g)) %>% 
  mutate(leaf_wet_mass = leaf_wet_mass_g - cup_mass) %>% 
  mutate(leaf_dry_mass = leaf_dry_mass - dry_cup_mass) %>% 
  select(-leaf_wet_mass_g) %>% 
  mutate(leaf_lfm = 100*(leaf_wet_mass - leaf_dry_mass)/leaf_dry_mass)

stem_data <- stem_data %>% 
  mutate(stem_wet_mass = stem_wet_mass - cup_mass) %>% 
  mutate(stem_dry_mass = stem_dry_mass - dry_cup_mass) %>% 
  mutate(stem_lfm = 100*(stem_wet_mass - stem_dry_mass)/stem_dry_mass)
```

# Data Wrangling
## Stem Data
```{r}
stem_data <- stem_data %>% 
  mutate(stem_width = (stem_width1 + stem_width2)/2) #average of two stem width measurements taken

leaf_area <- leaf_area %>% 
  mutate(plant = case_when(
    plant_number == 1 ~ 'P1',
    plant_number == 2 ~ 'P2',
    plant_number == 3 ~ 'P3',
    plant_number == 4 ~ 'P4',
    plant_number == 5 ~ 'P5',
    plant_number == 6 ~ 'P6',
    plant_number == 7 ~ 'P7',
    plant_number == 8 ~ 'P8',
    plant_number == 9 ~ 'P9',
    plant_number == 10 ~ 'P10',
    plant_number == 11 ~ 'P11',
    plant_number == 12 ~ 'P12'
  ))
```

### Stem SA:V
For the stem, using the formula for a frustrum where the two stem diameters are used to calculate the radii of each side of the frustrum, and each stem segment was cut to 5cm long; note that stem width is measured in mm while height is measured in cm; since we want it to all be in cm, we adjust widths acccordingly
```{r}
stem_data <- stem_data %>% 
  mutate(stem_width2 = ifelse(stem_width2 > 10, stem_width2/100, stem_width2)) %>% # some outliers where things look like measurements were entered incorrectly; adjusting them so they make sense
  mutate(stem_width1 = ifelse(stem_width1 > 10, stem_width1/10, stem_width1)) %>% 
  mutate(stem_sa = (pi*(0.1*stem_width1+0.1*stem_width2))*sqrt((0.1*stem_width1+0.1*stem_width2)^2 + 5^2)) %>% 
  mutate(stem_vol = (pi*5/3*((0.1*stem_width1)^2+(0.1*stem_width1*0.1*stem_width2)+(0.1*stem_width2)^2))) %>% 
  mutate(stem_sav = stem_sa/stem_vol)
```

## VOC Data
We're interested in the area under the curve of the VOCs over time plot. To calculate this, I used the function below to approximate the area under the curve using the trapezoidal rule:
```{r}
voc_data <- voc_data %>% 
  mutate(species = case_when(
    species == "Iris douglasiana" ~ "IRIDOU",
    species == "Erigeron karvinskianus" ~ "ERIKAR",
    species == "Salvia apiana" ~ "SALAPI",
    species == "Salvia leucophylla" ~ "SALLEU",
    species == "Artemesia californica" ~ "ARTCAL",
    species == "Ceanothus griseus" ~ "CEAGRI",
    species == "Malosma laurina" ~ "MALLAU",
    species == "Heteromeles arbutifolia" ~ "HETARB"
  )) %>% 
  group_by(leaf_mass) %>% 
  summarise(species = species, method = method,
         auc_voc = area_under_curve(time, voc), # Area under curve
         avg_voc = sum(voc)/10) # to get voc per minute average
```

## Averaging Traits
Averaging traits of interest so that there is one value per plant
```{r}
stem_data_avgs <- stem_data %>% 
  group_by(species, plant) %>% 
  summarise(stem_lfm = mean(stem_lfm), stem_width = mean(stem_width), stem_wet_mass = mean(stem_wet_mass), stem_dry_mass = mean(stem_dry_mass), stem_sav = mean(stem_sav))

leaf_data_avgs <- leaf_data %>% 
  group_by(species, plant) %>% 
  summarise(leaf_lfm = mean(leaf_lfm), thickness = mean(thickness), leaf_wet_mass = mean(leaf_wet_mass), leaf_dry_mass = mean(leaf_dry_mass))

leaf_area_avgs <- leaf_area %>% 
  group_by(species, plant) %>% 
  summarise(leaf_area = mean(area_cm2))

voc_data_avgs <- voc_data %>% 
  group_by(species) %>% 
  summarise(auc_voc = mean(auc_voc), avg_voc = mean(avg_voc), norm_avg_voc = mean(avg_voc)/mean(leaf_mass), norm_auc_voc = mean(auc_voc)/mean(leaf_mass))
```

## Combining Traits Data
Combining datasets so that they can be easily integrated into the flamm data
- First leaf_data and stem_data
```{r}
## Adding IRIDOU to stem_data w/ NAs
species <- c(rep('IRIDOU', 9))
plant <- c('P1', 'P2', 'P3', 'P4', 'P5', 'P6', 'P7', 'P8', 'P9')
na <- c(rep(999999, 9)) # so that IRIDOU data does not get lost in the merge
stem_data_avgs_iridou <- data.frame(species, plant, na, na, na, na, na)
colnames(stem_data_avgs_iridou) <- colnames(stem_data_avgs)
stem_data_avgs <- rbind(stem_data_avgs, stem_data_avgs_iridou)
## Combining
leaf_stem_df <- merge(leaf_data_avgs, stem_data_avgs, by = c('species', 'plant'))
```

### Calculating total LFM
```{r}
leaf_stem_df <- leaf_stem_df %>% 
  mutate(lfm = ifelse(stem_wet_mass > 99999, leaf_lfm, 100*((stem_wet_mass + leaf_wet_mass) - (stem_dry_mass + leaf_dry_mass))/(stem_dry_mass + leaf_dry_mass)))

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  #filter(lfm > 0  & lfm < 500) %>% 
  ggplot(aes(x = lfm, y = stem_lfm)) +
    labs(x = 'LFM', y = 'LFM (stem)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  #filter(lfm > 0  & lfm < 500) %>% 
  #filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  ggplot(aes(x = lfm, y = leaf_lfm)) +
    labs(x = 'LFM', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  #filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  #filter(stem_lfm > 0  & stem_lfm < 600) %>% 
  ggplot(aes(x = stem_lfm, y = leaf_lfm)) +
    labs(x = 'LFM (stem)', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()
```

### Dry Matter Content
```{r}
leaf_stem_df <- leaf_stem_df %>% 
  mutate(leaf_dmc = leaf_dry_mass/leaf_wet_mass) %>% 
  mutate(stem_dmc = ifelse(stem_wet_mass > 99999, 999999, stem_dry_mass/stem_wet_mass)) %>% 
  mutate(dmc = ifelse(stem_wet_mass > 99999, leaf_dry_mass/leaf_wet_mass,
                      (leaf_dry_mass + stem_dry_mass/leaf_wet_mass + stem_wet_mass)))

leaf_stem_df %>% 
  filter(species != 'QUEAGR' & species != 'EUGL') %>% 
  filter(stem_wet_mass < 99999) %>% 
  pivot_longer(cols = c(stem_dmc, dmc, leaf_dmc, stem_lfm, lfm), names_to = 'var', values_to = 'val') %>% 
  ggplot(aes(x = species, y = val)) +
    geom_boxplot() +
    facet_wrap(~var, scales = 'free') +
    theme_bw()

library(GGally) #ggcorr
leaf_stem_df %>% 
  filter(species != 'QUEAGR' & species != 'EUGL') %>% 
  filter(stem_wet_mass < 99999, leaf_lfm > 0) %>% 
  select(stem_dmc, dmc, leaf_dmc, stem_lfm, lfm, leaf_lfm) %>% 
  ggcorr(label = T)
```

- Now, stem_leaf_ratio
note: for now, not going to add or takeaway any data prior to merging, so we will lose some of the data from the above dataset. Hopefully, Zoe, Rob and I will figure this out by looking into where the missing trait data could be.
```{r}
stem_leaf_ratio_clean <- stem_leaf_ratio %>% 
  select(species, plant, date, total_branch_mass, total_leaf_mass) %>% 
  mutate(stem_mass_ratio = total_branch_mass/(total_branch_mass + total_leaf_mass)) %>% 
  mutate(leaf_mass_ratio = total_leaf_mass/(total_branch_mass + total_leaf_mass)) 
  
stem_leaf_ratio_clean <- subset(stem_leaf_ratio_clean, !(species == 'SALAPI' & plant %in% c('P9', 'P10') & date == '2022-08-24')) # Removing duplicated values for SALAPI P9 and P10 (must've accidentally measured it twice; I took the first measurements, as they were inputted by the same person as the other SALAPI measurements)

leaf_stem_v2_df <- merge(stem_leaf_ratio_clean, leaf_stem_df, by = c('species', 'plant'))
```

### Leaf Area and SA:V

- Leaf Area
```{r}
# Adding NA values for HETARB P10-12
species <- c(rep('HETARB', 3))
plant <- c('P10', 'P11', 'P12')
na <- c(rep(999999, 3)) # so that IRIDOU data does not get lost in the merge
leaf_area_avgs_hetarb <- data.frame(species, plant, na)
colnames(leaf_area_avgs_hetarb) <- colnames(leaf_area_avgs)
leaf_area_avgs <- rbind(leaf_area_avgs, leaf_area_avgs_hetarb)

main_traits_df <- merge(leaf_stem_v2_df, leaf_area_avgs, by = c('species', 'plant'))
```

- Leaf SA:V
Calculating surface area by multiplying leaf area x 2 and then volume multiplying leaf area times thickness
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(leaf_sav = ifelse(leaf_area > 99999, 999999, (2*leaf_area)/(thickness*leaf_area)))
```

### VOCs
```{r}
# Adding NA values for species we don't have VOC data for (MANDEN, QUEAGR, EUCGLO)
species <- c('MANDEN', 'QUEAGR', 'EUGL')
na <- c(rep(999999, 3)) # so that IRIDOU data does not get lost in the merge
voc_data_avgs_nas <- data.frame(species, na, na, na, na)
colnames(voc_data_avgs_nas) <- colnames(voc_data_avgs)
voc_data_avgs <- rbind(voc_data_avgs, voc_data_avgs_nas)

main_traits_df <- merge(main_traits_df, voc_data_avgs, by = 'species')
```

- Dealing with NAs from above (HETARB, IRIDOU)
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(stem_lfm = ifelse(stem_lfm > 99999, NA, stem_lfm)) %>% 
  mutate(stem_width = ifelse(stem_width > 99999, NA, stem_width)) %>%
  mutate(stem_wet_mass = ifelse(stem_wet_mass > 99999, NA, stem_wet_mass)) %>% 
  mutate(stem_dry_mass = ifelse(stem_dry_mass > 99999, NA, stem_dry_mass)) %>% 
  mutate(leaf_area = ifelse(leaf_area > 99999, NA, leaf_area)) %>% 
  mutate(auc_voc = ifelse(auc_voc == 999999, NA, auc_voc)) %>% 
  mutate(avg_voc = ifelse(avg_voc == 999999, NA, avg_voc)) %>% 
  mutate(norm_auc_voc = ifelse(norm_auc_voc == 999999, NA, norm_auc_voc)) %>% 
  mutate(norm_avg_voc = ifelse(norm_avg_voc == 999999, NA, norm_avg_voc)) %>% 
  mutate(stem_dmc = ifelse(stem_dmc == 999999, NA, stem_dmc)) %>% 
  mutate(stem_sav = ifelse(stem_dmc == 999999, NA, stem_sav)) %>% 
  mutate(leaf_sav = ifelse(leaf_sav == 999999, NA, leaf_sav))
```

## Changing 'plant' to numeric
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(plant = case_when(
    plant == 'P1' ~ 1,
    plant == 'P2' ~ 2,
    plant == 'P3' ~ 3,
    plant == 'P4' ~ 4,
    plant == 'P5' ~ 5,
    plant == 'P6' ~ 6,
    plant == 'P7' ~ 7,
    plant == 'P8' ~ 8,
    plant == 'P9' ~ 9,
    plant == 'P10' ~ 10,
    plant == 'P11' ~ 11,
    plant == 'P12' ~ 12
  )) %>% 
  mutate(species = ifelse(species == 'EUGL', 'EUCGLO', species))
```

## Calculating LMA, SLA
Note that for ERIKAR P4-P10, all ARTCAL, SALLEU P11, SALAPI P12, leaf dry mass was for all of the leaves rather than just the five used in the leaf scans, hence why those values are NA below
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(LMA = ifelse(species == 'ARTCAL' | species == 'ERIKAR' & plant %in% c(4:10), NA, leaf_dry_mass/leaf_area),
         SLA = ifelse(species == 'ARTCAL' | species == 'ERIKAR' & plant %in% c(4:10), NA, leaf_area/leaf_dry_mass))
```

Imputing species mean LMAs/SLAs for missing values (for HETARB and ERIKAR)
```{r}
# LMA
mean_HETARB_LMA <- main_traits_df %>% 
  drop_na(LMA) %>% 
  filter(species == 'HETARB') %>% 
  summarise(LMA = mean(LMA))

main_traits_df <- main_traits_df %>% 
  mutate(LMA = ifelse(species == 'HETARB' & is.na(LMA), mean_HETARB_LMA[1,1], LMA))

mean_ERIKAR_LMA <- main_traits_df %>% 
  drop_na(LMA) %>% 
  filter(species == 'ERIKAR') %>% 
  summarise(LMA = mean(LMA))

main_traits_df <- main_traits_df %>% 
  mutate(LMA = ifelse(species == 'ERIKAR' & is.na(LMA), mean_ERIKAR_LMA[1,1], LMA))

# SLA
mean_HETARB_SLA <- main_traits_df %>% 
  drop_na(SLA) %>% 
  filter(species == 'HETARB') %>% 
  summarise(SLA = mean(SLA))

main_traits_df <- main_traits_df %>% 
  mutate(SLA = ifelse(species == 'HETARB' & is.na(SLA), mean_HETARB_SLA[1,1], SLA))

mean_ERIKAR_SLA <- main_traits_df %>% 
  drop_na(SLA) %>% 
  filter(species == 'ERIKAR') %>% 
  summarise(SLA = mean(SLA))

main_traits_df <- main_traits_df %>% 
  mutate(SLA = ifelse(species == 'ERIKAR' & is.na(SLA), mean_ERIKAR_SLA[1,1], SLA))
```

### ARTCAL
For ARTCAL, we did not get the correct measurements for any plants, so we had to use a separate potted plant to get a species average and impute that value for the LMA for ARTCAL
```{r}
# load in data
artcal_lma <- read_csv(here('data', 'raw-data', 'plant-traits', 'ARTCAL_LMA.csv')) %>%
  clean_names() %>% 
  rename(cup_number = cup_no)

# subtracting cup weights
artcal_lma <- merge(artcal_lma, cup_weights_data, by = 'cup_number')
artcal_lma <- artcal_lma %>% 
  mutate(wet_mass = wet_mass - cup_mass,
         dry_mass = dry_mass - dry_cup_mass)

# calculating average lma
artcal_lma_means <- artcal_lma %>% 
  group_by(id) %>% 
  summarise(wet_mass = mean(wet_mass), dry_mass = mean(dry_mass), leaf_area = sum(area)) %>% 
  mutate(LMA = dry_mass/leaf_area, SLA = leaf_area/dry_mass)
mean_ARTCAL_LMA <- artcal_lma_means %>% 
  summarise(LMA = mean(LMA), SLA = mean(SLA))

# adding to main dataset
main_traits_df <- main_traits_df %>% 
  mutate(SLA = ifelse(species == 'ARTCAL' & is.na(SLA), mean_ARTCAL_LMA$SLA[1], SLA)) %>% 
  mutate(LMA = ifelse(species == 'ARTCAL' & is.na(LMA), mean_ARTCAL_LMA$LMA[1], LMA))
```

## Small Adjustments
```{r}
# Filtering out EUCGLO and QUEAGR; not including these species in this analysis
main_traits_df <- main_traits_df %>% 
  filter(species != 'EUCGLO',
         species != 'QUEAGR') %>% 
  mutate(species = ifelse(species == 'MANDEN', 'ARCDEN', species))
```

# Combining with Flam. Dataset
```{r}
flam.df <- read.csv(here('data', 'processed-data', 'clean_flamm_data.csv')) %>% 
  select(-date)

main_df <- merge(main_traits_df, flam.df, by = c('species', 'plant'))
```

## Small Adjustments (2)
```{r}
# One missing mpa value for HETARB; because we don't have many HETARB samples as is, let's impute the species mean for this mpa like we did for LMAs
mean_HETARB_mpa <- main_df %>% 
  drop_na(mpa) %>% 
  filter(species == 'HETARB') %>% 
  summarise(mpa = mean(mpa))

main_df <- main_df %>% 
  mutate(mpa = ifelse(species == 'HETARB' & is.na(mpa), mean_HETARB_mpa[1,1], mpa))

# Getting rid of duplicate rows
duplicates <- main_df %>% 
  unite('id', c(species, plant, rep), sep = '_', remove = F)

duplicates[duplicated(duplicates$id), ]
```

# Sample Weight/Size
## Dry Weight, Wet Weight
```{r}
main_df <- main_df %>% 
  mutate(wet_mass = ifelse(species == 'IRIDOU', leaf_wet_mass, # Because IRIDOU has no stems
                           stem_wet_mass + leaf_wet_mass),
         dry_mass = ifelse(species == 'IRIDOU', leaf_dry_mass, 
                           stem_dry_mass + leaf_dry_mass),
         gdw_gfw = dry_mass/(wet_mass+dry_mass),
         dw_flam_sample = gdw_gfw * sample_wt, # need to change to sample weight
         ww_flam_sample = sample_wt - dw_flam_sample)
```

## Sample Volume
```{r}
main_df <- main_df %>% 
  mutate(branch_volume = branch_length*branch_width*branch_height)
```

# Branching
For branching, Rob suggested we use a branching per sample length metric that gives us more information regarding the 'bushiness' of the sample rather than just the number of branches
```{r}
main_df <- main_df %>% 
  rename(no_branches = branching) %>%  # to rename the initial variable to reflect # of branches so that 'branching' reflects what we're actually interested in
  mutate(branching = (no_branches + 1)/branch_length)
```

# Density
Also following a suggestion from Rob, using sample density (g/cm3) as a comparable variable to fuel density which is of-concern on the landscape scale
```{r}
main_df <- main_df %>% 
  mutate(sample_density = sample_wt/branch_volume)
```

# Dealing w/ NAs
For some of the plant traits, there are NA values that are due to randomly missing data. For these, we'd like to keep the datapoints in the analysis, but -- to do so -- we'll have to impute the data by using either the median or the mean values. First, we should take a look at which values are NA that we could deal with
```{r}
summary(main_df)
```

For SALAPI P8 R2, we have NAs for branch dimensions, sample weight, branching and mpa; because this is only one datapoint, I think the best thing to do is to delete this datapoint
```{r}
main_df <- main_df %>% 
  drop_na(branch_height)
```

# Dry Weight (Deviation from Species Mean)
```{r}
dw_avgs <- main_df %>% 
  drop_na(dw_flam_sample) %>% 
  group_by(species) %>% 
  summarise(dw_mean = mean(dw_flam_sample), dw_sd = sd(dw_flam_sample))

main_df <- main_df %>% 
  mutate(dw_sppdev = ifelse(is.na(dw_flam_sample), NA, case_when(
    species == 'ARCDEN' ~ (dw_flam_sample - as.numeric(dw_avgs[1,2]))/as.numeric(dw_avgs[1,3]),
    species == 'ARTCAL' ~ (dw_flam_sample - as.numeric(dw_avgs[2,2]))/as.numeric(dw_avgs[2,3]),
    species == 'CEAGRI' ~ (dw_flam_sample - as.numeric(dw_avgs[3,2]))/as.numeric(dw_avgs[3,3]),
    species == 'ERIKAR' ~ (dw_flam_sample - as.numeric(dw_avgs[4,2]))/as.numeric(dw_avgs[4,3]),
    species == 'HETARB' ~ (dw_flam_sample - as.numeric(dw_avgs[5,2]))/as.numeric(dw_avgs[5,3]),
    species == 'IRIDOU' ~ (dw_flam_sample - as.numeric(dw_avgs[6,2]))/as.numeric(dw_avgs[6,3]),
    species == 'MALLAU' ~ (dw_flam_sample - as.numeric(dw_avgs[7,2]))/as.numeric(dw_avgs[7,3]),
    species == 'SALAPI' ~ (dw_flam_sample - as.numeric(dw_avgs[8,2]))/as.numeric(dw_avgs[8,3]),
    species == 'SALLEU' ~ (dw_flam_sample - as.numeric(dw_avgs[9,2]))/as.numeric(dw_avgs[9,3]))))
```

# Cutting IRIDOU
```{r}
main_df <- main_df %>% 
  filter(species != 'IRIDOU')
```


# Writing csv
```{r}
summary(main_traits_df)
write.csv(main_traits_df, here('data', 'processed-data', 'plant_traits_data.csv'))

summary(main_df)
write.csv(main_df, here('data', 'processed-data', 'main_dataset.csv'))
```
