---
title: "Wonderlabs"
author: "Indra Boving"
date: "11/24/2021"
output: html_document
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
here = here::here
source(here("scripts", "source_code.R"))

# ggplot theme for script (Indra)
ggplot <- function(...) { ggplot2::ggplot(...) + 
    theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
          panel.grid.major = element_blank(),  # Hide major gridlines
          panel.grid.minor = element_blank()) # Hide minor gridlines
}
```

# Raw Data
```{r, warning = FALSE}
data <- read_csv(here("data", "raw-data", "survey", "wonderlabs.csv"), show_col_types = FALSE) %>% 
  clean_names() %>% 
  filter(response_type != "Survey Preview") %>% 
  dplyr::select(-progress, -finished, -distribution_channel) %>% 
  separate(control, c("rent_own", "control"), sep = ",") %>% 
  separate(maintenance, c("maint1", "maint2", "maint3", "maint4", "maint5"), sep = ",") %>% 
  separate(time_30mins, c("time1", "time2"), sep = ",") %>% 
  mutate(birthyear = case_when(
    birthyear == "12/31/54" ~ 1954, 
    TRUE ~ as.numeric(birthyear))) %>% 
  mutate(income = as.numeric(income)) %>% 
  mutate(age = 2021 - birthyear) 
```

#-----------------------------

# Demographics

## Data Wrangling
```{r, warning = FALSE}
demographics <- data %>% 
  dplyr::select(response_id, user_language, rent_own, control, birthyear, gender, race, language, income, age, time1)

income <- demographics %>% 
  drop_na(income) %>% 
  summarise(mean = mean(income), 
            median = median(income), 
            range = range(income))
income

race<- demographics %>% 
  drop_na(race) %>% 
  tabyl(race)
race
```

## Rent vs. Own
```{r}
renters_counts <- demographics %>% 
  drop_na(rent_own, control) %>% 
  tabyl(rent_own, control)
renters_counts

renters_counts_props <- renters_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
renters_counts_props

kable(renters_counts_props) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

## First Language
```{r, include = FALSE}
language_counts <- demographics %>% 
  drop_na(language) %>% 
  tabyl(language)
language_counts
```

## Age
```{r}
age_table <- demographics %>% 
  drop_na(age) %>% 
  filter(age > 0) %>% 
  summarise(mean = mean(age), 
            range = range(age), 
            median = median(age))
age_table
```

## Gender
```{r}
gender_table <- demographics %>% 
  drop_na(gender) %>% 
  tabyl(gender)
gender_table
```


#-----------------------------

# Ecosystem Services

```{r}
es_data_id <- data %>% 
  #dplyr::select(3, 5:28) %>%  #includes response ID
  dplyr::select(3, 5:28) %>% 
  drop_na %>% 
  #as.data.frame() %>% 
  mutate_all(funs(str_replace(., "No Importance", "1"))) %>% 
  mutate_all(funs(str_replace(., "Low Importance", "2"))) %>% 
  mutate_all(funs(str_replace(., "Medium Importance", "3"))) %>% 
  mutate_all(funs(str_replace(., "Very High Importance", "5"))) %>%   mutate_all(funs(str_replace(., "High Importance", "4"))) %>%  
  mutate_at((2:25), as.numeric) %>% 
  pivot_longer(!response_id, names_to = "item", values_to = "count") %>% 
  mutate(category = case_when(
    item == "food_or_home_garden" ~ "provisioning",
    item == "commercial_agriculture" ~ "provisioning",
    item == "medicinal" ~ "provisioning",
    item == "edible_wild_plants" ~ "provisioning",
    item == "food_or_habitat_for_wild_animals" ~ "provisioning",
    item == "food_or_space_for_pet_animals" ~ "provisioning",
    item == "air_quality_control" ~ "regulating",
    item == "air_quality_control" ~ "regulating",
    item == "erosion_control" ~ "regulating",
    item == "pollination" ~ "regulating",
    item == "cooling_provides_shade" ~ "regulating",
    item == "soil_fertility_control" ~ "regulating",
    item == "waste_treatment" ~ "regulating",
    item == "protect_or_shield_my_home_from_wildfire" ~ "regulating",
    item == "place_for_private_recreation_physical_activity" ~ "cultural",
    item == "place_for_socializationl" ~ "cultural",
    item == "privacy" ~ "cultural",
    item == "religious_or_spiritual" ~ "cultural",
    item == "beauty" ~ "cultural",
    item == "firewood" ~ "provisioning",
    item == "sense_of_place" ~ "cultural",
    item == "tranquility_mental_health" ~ "cultural",
    item == "inspiration_for_art_design" ~ "cultural",
    item == "observing_the_natural_world" ~ "cultural",
    item == "block_or_deter_entry_from_my_home_or_property" ~ "cultural", 
item == "place_for_socialization" ~ "cultural")) %>% 
  mutate(category_n = case_when(
    category == "provisioning" ~ 8, 
    category == "cultural" ~ 11,
    category == "regulating" ~ 8))


#summary table: 
totals_by_id <- es_data_id %>% 
  group_by(response_id, category) %>% 
  mutate(sum_category = sum(count)) %>% 
  mutate(score = sum_category/category_n) %>%
ungroup() %>%
group_by(category) %>%
mutate(mean_score = mean(score)) %>% 
ungroup() %>% 
group_by(item) %>% 
mutate(sum_item = sum(count)) %>% 
add_tally() %>% 
mutate(score_item = sum_item/n) %>% 
mutate(mean_score_item = mean(score_item))
```

### ES vs. Yard Maintanence

```{r, warning = FALSE}
maintainance_1 <- data %>% 
  dplyr::select(response_id, maint1, maint2, maint3, trait, writein_trait, rent_own, control, time1) %>% 
  drop_na(maint1)
```

```{r}
maintainance <- merge(totals_by_id, maintainance_1, by = "response_id")  
```

```{r}
maintainance %>% 
  #filter(category == "cultural") %>% 
  dplyr::group_by(item, time1) %>% 
  summarise(mean_score_item_time1 = mean(score_item), time1 = time1, item = item, category = category) %>% 
  ggplot(aes(y = mean_score_item_time1, x = time1, fill = item)) +
  geom_col(position = "dodge") +
  facet_wrap(~category, nrow = 3) +
  labs(y = "Score", x = "Importance") +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
maintainance %>%
  group_by(time1, category) %>% 
  mutate(mean_score_time = mean(score)) %>% 
ggplot(aes(y = mean_score_time, x = time1, fill = category)) +
  geom_col(position = "dodge") +
  theme_bw() +
  scale_fill_brewer(palette = "YlOrRd") +
  labs(y = "Score", x = "Importance") +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
maintainance %>%
  ungroup() %>% 
  mutate(trait_names = case_when(
    trait == "Produces food or edibles" ~ "Edible", 
    trait == "Fragrant (smells good)" ~ "Fragrant", 
    trait == "Requires little watering to stay healthy" ~ "Drought tolorant", 
    trait  == "Flowers" ~ "Flowers", 
    writein_trait %in% c("wildlife value and screening", "wildlife habitat", "all of the above, especially wildlife food", "good for honey bees & Humming birds", "Hosts birds, provides shade","Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.") ~ "Wildlife", 
    writein_trait %in% c("Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.", "all of the above") ~ "Flowers", 
    writein_trait %in% c("Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.", "all of the above") ~ "Drought tolorant"
  )) %>% 
  group_by(item, trait) %>% 
  mutate(mean_score_trait = mean(score)) %>% 
  filter(trait != "Other:") %>% 
  drop_na(item) %>% 
ggplot(aes(y = mean_score_trait, x = trait_names)) +
  geom_col(position = "dodge") +
  theme_bw() +
  scale_fill_viridis_d() +
  labs(y = "Score", x = "Trait") +
  theme(legend.position = "bottom") +
  facet_wrap(~category)
```

# Plant Traits
```{r}
plants <- data %>% 
  dplyr::select(response_id, trait, writein_trait, time1) %>% 
 mutate(trait_names = case_when(
    trait == "Produces food or edibles" ~ "Edible", 
    trait == "Fragrant (smells good)" ~ "Fragrant", 
    trait == "Requires little watering to stay healthy" ~ "Drought Tolerant", 
    trait  == "Flowers" ~ "Flowers", 
    writein_trait %in% c("wildlife value and screening", "wildlife habitat", "all of the above, especially wildlife food", "good for honey bees & Humming birds", "Hosts birds, provides shade","Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.") ~ "Wildlife Habitat", 
    writein_trait %in% c("Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.", "all of the above") ~ "Flowers", 
    writein_trait %in% c("Natives that are food or shelter to native insects. Plants that require a small amount of water. I also like flowering plants that smell nice.", "all of the above") ~ "Drought Tolerant"
  )) %>% 
  drop_na(trait_names)
```

```{r}
pct_plants <- plants %>% 
  group_by(trait_names) %>% 
  tally()
pct_plants <- pct_plants %>% 
  mutate(pct = n/sum(pct_plants$n)*100)
  
plants_text <- data.frame(trait_names = c('Wildlife Habitat', 'Fragrant', 'Flowers', 'Edible', 'Drought Tolerant'),
                          trait_labs = c('Wildlife Habitat (C,R,P)', 'Fragrant (C,R)', 'Flowers (C,R)', 'Edible (P)', 'Drought Tolerant (C,R)'),
                          y.lab = c(26, 13, 17, 29.5, 40),
                          lab.color = c(rep('black', 4), 'white'))

traits.survey.plot <- pct_plants %>% 
  ggplot(aes(x = trait_names, y = pct)) +
  geom_bar(fill = 'gray30', stat = 'identity') +
  geom_text(data = plants_text, aes(x = trait_names, label = trait_labs, y = y.lab, color = lab.color),
            fontface = 'bold', size = 7.5) +
  scale_color_manual(values = c('black', 'white')) +
  labs(y = '% of Responses', 
       x = "Trait",
       fill = "Trait") +
  coord_flip() +
  scale_x_discrete(position = 'top') +
  theme(legend.position = 'none',
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 20, face = 'bold'),
        axis.text.x = element_text(size = 18),
        #axis.text.y = element_text(size = 14, face = 'bold'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin = unit(c(.25, .15, 1.46, .15), "cm"))
traits.survey.plot
```

Setting up dataframe showing how our selected speccies fit into the categories described in the above barplot
```{r}
species <- c('ARCDEN', 'ARTCAL', 'CEAGRI', 'ERIKAR', 'HETARB', 'IRIDOU', 'MALLAU', 'SALAPI', 'SALLEU')
`Drought Tolerant` <- c(1, 1, 1, 0.75, 1, 0.5, 1, 1, 1)
Edible <- c(0.5, 0.5, 0.5, 1, 0.5, 0, 0.5, 1, 1)
`Wildlife Habitat` <- c(1, 1, 1, 0.5, 1, 0.5, 1, 1, 1)
Flowers <- c(1, 0.5, 1, 1, 1, 1, 0.5, 1, 1)
Fragrant <- c(0, 1, 1, 0, 0.5, 0, 1, 1, 1)
Flammable <- c(0.4287465, 0.3767518, 0.6809514, 0.1153639, 0.8574929, 0, 0.7776964, 1, 0.8411896)
species_traits <- data.frame(species, `Drought Tolerant`, Edible, `Wildlife Habitat`, Flowers, Fragrant, Flammable) %>% 
  pivot_longer(cols = !c(species), names_to = 'trait', values_to = 'rep')

repr.plot <- ggplot(species_traits, aes(x = trait, y = species, fill= rep)) + 
  geom_tile() +
  scale_fill_gradient(low="gray90", high="gray30") +
  labs(fill = 'Representative?') +
  theme(panel.background = element_rect(fill='white', colour='black'),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title = element_blank(),
          axis.text = element_text(size = 12, face = 'bold'),
          legend.title = element_text(size = 12, face = 'bold'),
          legend.position = 'bottom')
repr.plot
```

## Arranged
```{r}
ggarrange(traits.survey.plot, repr.plot,
          labels = c('B', 'C'),
          font.label = list(size = 28),
          ncol = 1)
#ggsave(here('figures', 'main_figures', 'desirable.traits.survey.png'), height = 12, width = 10)
```

# Plant Species List
```{r}
species_list_data <- read.csv(here('data', 'processed-data', 'species_list_data.csv'))[1:145,]

species_list_data <- species_list_data %>% 
  mutate(selected = ifelse(
    species.name %in% c('Artemesia californica', 'Ceanothus sp.', 'Erigeron karvinskianus', 'Heteromeles arbutifolia', 'Iris sp.', 'Arctostaphylos sp.', 'Salvia sp.'), 'Yes', 'No'))

species_list_summary <- species_list_data %>% 
  drop_na(species.name) %>% 
  group_by(species.name, selected) %>% 
  tally() %>% 
  mutate(species.name = as.factor(species.name)) %>% 
  arrange(n)

# adding in malosma laurina with # of responses = 0 to summary
mallau_add <- data.frame(species.name = 'Malosma laurina', selected = 'Yes', n = 0)
species_list_summary <- rbind(species_list_summary, mallau_add)

species_list_summary$species.name <- factor(species_list_summary$species.name, levels = species_list_summary$species.name[order(species_list_summary$n)])
species_list_summary$selected <- factor(species_list_summary$selected, levels = c('Yes', 'No'))


species_list <- species_list_summary %>% 
ggplot(aes(x = n, y = species.name, color = selected)) +
  geom_point(size = 4) +
  geom_segment(aes(x=0, xend=n, y=species.name, yend=species.name), size = 1) +
  scale_color_manual(values = c('black', scales::alpha('gray70', 0.7))) +
  labs(color = 'Tested?', x = '# of Responses') +
  theme(legend.position = c(0.65, 0.1),
        axis.title.y = element_blank(),
        axis.title.x = element_text(face = 'bold', size = 16),
        legend.title = element_text(face = 'bold', size = 14),
        legend.text = element_text(size = 12),
        axis.text.y = element_text(face = 'italic', size = 14),
        axis.text.x = element_text(size = 14))

ggarrange(species_list,
          font.label = list(size = 28),
          ncol = 1)

ggsave(here('figures', 'supp_figures', 'species.tested.vs.survey.png'), height = 15, width = 7)
```

-----------------------------------------------------------------
```{r}
es_data <- data %>% 
  #select(3, 5:28) %>%  #includes response ID
  dplyr::select(5:28) %>% 
  drop_na()

options(digits = 2)

make_p_tab <- function(x) {
  es_data %>% 
    drop_na() %>% 
  dplyr::select(x) %>% 
  tabyl(x)
}

to_do <- colnames(es_data)

dt_list <- map(to_do, make_p_tab)

mergeDTs <- function(dt_list, by = NULL, sort = FALSE) {
  Reduce(
    function(...) {
      merge(..., by = by, all = TRUE, sort = sort)
    }, dt_list)
}
```

#Likert Scale

```{r}
es_data <- data %>% 
  dplyr::select(5:28) %>% 
  drop_na %>% 
  as.data.frame() %>% 
  mutate_if(is.character, as.factor)

class(es_data$medicinal)

str(es_data)

#diagnose(es_data)

mylevels <- c('No Importance', 'Low Importance', 'Medium Importance', 'High Importance', 'Very High Importance')

for(i in seq_along(es_data)) {
	es_data[,i] <- factor(es_data[,i], levels=mylevels)
}

sapply(es_data, class) #Verify that all the columns are indeed factors
sapply(es_data, function(x) { length(levels(x)) } ) # The number of levels in each factor

summary <- summary(es_data)

es_likert <- likert::likert(es_data)
es_likert

es_likert_counts <- summary(es_likert)

plot(es_likert)

likert.heat.plot(es_likert)

```

```{r}
#compare different service types: 
es_likert_df <- as.data.frame(es_likert$results)

es_likert_df <- es_likert_df %>% 
  mutate(category = case_when(
    Item == "food_or_home_garden" ~ "provisioning",
    Item == "commercial_agriculture" ~ "provisioning",
    Item == "medicinal" ~ "provisioning",
    Item == "edible_wild_plants" ~ "provisioning",
    Item == "food_or_habitat_for_wild_animals" ~ "provisioning",
    Item == "food_or_space_for_pet_animals" ~ "provisioning",
    Item == "air_quality_control" ~ "regulating",
    Item == "air_quality_control" ~ "regulating",
    Item == "erosion_control" ~ "regulating",
    Item == "pollination" ~ "regulating",
    Item == "cooling_provides_shade" ~ "regulating",
    Item == "soil_fertility_control" ~ "regulating",
    Item == "waste_treatment" ~ "regulating",
    Item == "protect_or_shield_my_home_from_wildfire" ~ "regulating",
    Item == "place_for_private_recreation_physical_activity" ~ "cultural",
    Item == "place_for_socializationl" ~ "cultural",
    Item == "privacy" ~ "cultural",
    Item == "religious_or_spiritual" ~ "cultural",
    Item == "beauty" ~ "cultural",
    Item == "firewood" ~ "provisioning",
    Item == "sense_of_place" ~ "cultural",
    Item == "tranquility_mental_health" ~ "cultural",
    Item == "inspiration_for_art_design" ~ "cultural",
    Item == "observing_the_natural_world" ~ "cultural",
    Item == "block_or_deter_entry_from_my_home_or_property" ~ "cultural", 
Item == "place_for_socialization" ~ "cultural"))

  
es_likert_cat <- merge(es_likert_counts, es_likert_df) %>% 
  rename(low_percent = low, 
         neut_percent = neutral,
         high_percent = high, 
         mean_value = mean)
```

## Main Likert Scale Figure
```{r}
melted <- reshape::melt(es_likert_df) %>% 
  mutate(item_title = str_replace_all(Item, "_", " ")) %>% 
  mutate(item_title = str_to_title(item_title)) %>% 
  mutate(item_title = str_replace_all(item_title, " Or ", "/")) %>% 
  mutate(item_title = str_replace_all(item_title, "Private Recreation Physical Activity", "Private Recreation"),
         item_title = str_replace_all(item_title, "Of", "of"),
         item_title = str_replace_all(item_title, "For", "for"),
         item_title = str_replace_all(item_title, "The", "the"),
         item_title = str_replace_all(item_title, "My", "my"),
         item_title = str_replace_all(item_title, "From", "from"),
         item_title = str_replace_all(item_title, "Food/Home Garden", "Home Garden"),
         item_title = str_replace_all(item_title, "Cooling Provides Shade",
                                      "Cooling/Provides Shade"),
         item_title = str_replace_all(item_title, "Protect/Shield my Home from Wildfire",
                                      "Protect my Home from Wildfire")) %>%  # fixing some of the item titles
  filter(Item %in% c('commercial_agriculture', 'cooling_provides_shade', 'edible_wild_plants', 'erosion_control', 'food_or_habitat_for_wild_animals', 'food_or_home_garden', 'observing_the_natural_world', 'privacy', 'place_for_socialization', 'pollination', 'protect_or_shield_my_home_from_wildfire', 'religious_or_spiritual', 'soil_fertility_control')) %>% # selecting certain categories to include in figure
  mutate(cat_num = case_when(
    category %in% "provisioning" ~ 1, 
    category %in% "regulating" ~ 2, 
    category %in% "cultural" ~ 3, 
  ), 
  cat_num = as.numeric(cat_num))
melted

max_value_very_high <- melted %>% 
  filter(variable %in% c("Very High Importance")) %>% 
  group_by(item_title) %>% 
  mutate(very_high_value = max(value)) %>% 
  dplyr::select("item_title", "very_high_value")

melted <- merge(melted, max_value_very_high, by = c("item_title"))

leg <-  melted %>% 
  mutate(variable = case_when(variable == 'No Importance' ~ 'None',
                              variable == 'Low Importance' ~ 'Low',
                              variable == 'Medium Importance' ~ 'Medium',
                              variable == 'High Importance' ~ 'High',
                              variable == 'Very High Importance' ~ 'Very High')) %>% 
  mutate(variable = fct_relevel(variable, c('None', 'Low', 'Medium', 'High', 'Very High'))) %>% 
  #filter(cat_num == 1) %>% 
  ggplot()+
  geom_bar(aes(x = reorder(item_title,cat_num), y=value, fill=variable), position="stack", stat="identity")+
  coord_flip() + 
  labs(y = "% of Responses", x = '', fill = 'Importance')+
  xlab("")+
  scale_fill_manual(values = c("#595D7C", "#797F91", "#ADB6B2", "#C17C79", "#D24D4B"), guide = guide_legend(reverse = TRUE)) +
  #scale_fill_brewer(palette="BrBG", guide = guide_legend(reverse = TRUE))+
  theme(legend.position="bottom", 
        legend.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        legend.title = element_text(face = 'bold', size = 22))
        
leg

cultural <-  melted %>% 
  filter(cat_num == 3) %>% 
  ggplot()+
  geom_bar(aes(x = reorder(item_title, very_high_value), y=value, fill=variable), position="stack", stat="identity")+
  coord_flip() + 
  labs(x = "Cultural") + # change to title = 'cultural' to revert to previous aesthetic (same goes for below)
  scale_fill_manual(values = c("#595D7C", "#797F91", "#ADB6B2", "#C17C79", "#D24D4B")) +
  #scale_fill_brewer(palette="BrBG")+
  theme(legend.position="none", 
        axis.title.y = element_text(face = 'bold', size = 22, vjust = 4, hjust = 0.3),
        axis.text.y = element_text(size = 18),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        #plot.title = element_text(face = 'bold', size = 16),
        plot.margin = unit(c(.25, 0, .1, 0.2), "cm")
        )
cultural

reg <-  melted %>% 
  filter(cat_num == 2) %>% 
  ggplot()+
  geom_bar(aes(x = reorder(item_title, very_high_value), y=value, fill=variable), position="stack", stat="identity")+
  coord_flip() + 
  labs(x = "Regulating") +
  scale_fill_manual(values = c("#595D7C", "#797F91", "#ADB6B2", "#C17C79", "#D24D4B")) +
  #scale_fill_brewer(palette="BrBG")+
  theme(legend.position="none", 
        axis.title.y = element_text(face = 'bold', size = 22, vjust = 1),
        axis.text.y = element_text(size = 18),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        #plot.title = element_text(face = 'bold', size = 16),
        plot.margin = unit(c(0, 0, .1, .2), "cm")
        )
reg

prov <-  melted %>% 
  filter(cat_num == 1) %>% 
  ggplot()+
  geom_bar(aes(x = reorder(item_title, very_high_value), y=value, fill=variable), position="stack", stat="identity")+
  scale_y_continuous(limits = c(0,100)) +
  coord_flip() + 
  labs(x = "Provisioning", y = "% of Responses") +
  scale_fill_manual(values = c("#595D7C", "#797F91", "#ADB6B2", "#C17C79", "#D24D4B")) +
  #scale_fill_brewer(palette="BrBG")+
  theme(legend.position="none", 
        axis.title.y = element_text(face = 'bold', size = 22, vjust = 2.8),
        axis.text.y = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.title.x = element_text(face = 'bold', size = 20),
        #plot.title = element_text(face = 'bold', size = 16),
        plot.margin = unit(c(0, 0, .1, 0.2), "cm")
        )
prov

legend <- cowplot::get_legend(leg)
#scale_x_discrete(limits=c("StronglyAgree", "Agree", "DontKnow","Disagree","StronglyDisagree"))

survey_likert <- cowplot::plot_grid(cultural, reg, prov, legend, nrow = 4, align = "hv", axis = "lr", rel_heights = c(1, 1.2, 1.25, .25))
survey_likert

#ggsave(here("figures", 'main_figures', 'survey_likert.jpg'), height = 8, width = 12) 
```

# Main Figure: Likert and Desirable Traits
```{r}
plot_grid(survey_likert, traits.survey.plot,
          ncol = 2, nrow = 1, rel_widths = c(1.05, 0.7), labels = c('a', 'b'), label_size = 28)
ggsave(here("figures", 'main_figures', 'survey_figure.jpg'), height = 8, width = 17)
```

#-----------------------------

# Extra Figures

```{r}
melted %>% 
  filter(Item == "protect_or_shield_my_home_from_wildfire") %>% 
  filter(variable == "Very High Importance" | 
           variable == "High Importance") %>% 
  mutate(importance_prop = sum(value))
```

```{r}
melted %>% 
ggplot(aes(x = variable, y = value, fill = category)) +
         geom_col() +
  theme_bw() +
  scale_fill_brewer(palette = "PuBuGn") +
  labs(y = "Cumulative Percentage", x = "Importance") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
melted_means <- es_likert_cat %>% 
  dplyr::select(category, mean_value) %>% 
  reshape::melt()

melted_means %>% 
  ggplot(aes(y = value, x = category)) +
  #geom_point(aes(color = category))+
  # geom_boxplot(fill = NA, width = 0.2, outlier.color = NA, aes(color = category)) +
  geom_boxplot(aes(fill = category)) +
  scale_fill_brewer(palette="PuBuGn") +
  theme(legend.position ="none") +
  theme_bw() +
  labs(y = "Combined Score (1-5)", 
       x = "Category", 
       title = "Importance of ES Categories")
```

```{r}
melted_provisioning <- melted %>% 
  filter(category == "provisioning")
# %>% 
#   group_by(Item) %>% 
#   filter(variable == "Very High Importance") %>% 
#   mutate(max_value = max(value))
# 
# new <- merge(melted, melted_provisioning)

provisioning <- ggplot()+
  geom_bar(data = melted_provisioning, aes(x = item_title, y=value, fill=variable), position="stack", stat="identity", show.legend = FALSE)+
  coord_flip() + 
  ggtitle("Provisioning")+
  ylab("Percentage")+
  xlab("Type")+
  scale_fill_brewer(palette="YlOrRd") +
  theme(legend.title = element_blank()) +
  theme_bw()
```

```{r}
melted_cult <- melted %>% 
  filter(category == "cultural")

cultural <- ggplot()+
  geom_bar(data = melted_cult, aes(x = item_title, y= value, fill= variable), position="stack", stat="identity", show.legend = FALSE)+
  coord_flip() + 
  ggtitle("Cultural")+
  ylab("Percentage")+
  xlab("Type")+
  scale_fill_brewer(palette="YlOrRd") +
  theme(legend.title = element_blank()) +
  theme_bw()
cultural
```

```{r}
melted_reg <- melted %>% 
  filter(category == "regulating")

regulating <- ggplot() +
  geom_bar(data = melted_reg, aes(x = item_title, y=value, fill= variable), position="stack", stat="identity")+
  coord_flip() + 
  ggtitle("Regulating")+
  ylab("Percentage")+
  xlab("Type")+
  scale_fill_brewer(palette="YlOrRd") +
  theme(legend.position="bottom") +
  theme()
regulating
```

```{r}
melted <- melted %>% 
  mutate(Item = reorder(Item, category)) %>% 
  mutate(Item = reorder(Item, value))

ggplot()+
  geom_bar(data = melted, aes(x = item_title, y = value, fill= variable), position="stack", stat="identity")+
  coord_flip() + 
  #ggtitle("Cultural")+
  ylab("Percent")+
  xlab("Service")+
  scale_fill_brewer(palette="YlOrRd") +
  theme(legend.position = "bottom") +
  theme_bw() +
  facet_wrap(~category, nrow = 3, scales = "free")
```

```{r}
all <- # Create Flow timeseries plot that spans the grid by making one plot_grid
#   and then nest it inside of a second. Also, include a title at the top
#   for the whole figure.
  
title <- ggdraw() + draw_label("ES Categories and Valuation", fontface='bold')

plot_grid(provisioning, cultural, regulating, nrow = 4, c("", "", ""),  rel_heights = c(1, 1, 1))
```

